---
title: "Class Predictions"
output: html_document
date: "2023-03-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r echo=F, include=F}
library(tidyverse)
library(targets)
library(gghdx)
library(lubridate)
library(sysfonts)

font_add_google("Source Sans Pro")
showtext::showtext_auto()

target_ready <- c(T, F)[2]
if (target_ready) {
  tar_load(p_all_sites_events_pred_classifications)
}
if (!target_ready) {
  tar_source()
  tar_load(rainfall_impact_tbl)
  p_all_sites_events_pred_classifications <- plot_performance_all_sites(
    df = rainfall_impact_tbl,
    x = "precip_roll10",
    event = "fevent",
    thresh = 25,
    day_window = 60
  )
}

# unique(rainfall_impact_tbl$site_id)[5]
#
# plot_site_events_classified(df = rainfall_impact_tbl %>%
#                                 filter(site_id=="YE1116_0037"),
#                             x="precip_roll10",
#                             event = "fevent",
#                             thresh=25,
#                             day_window=60,plot_title = "boom")
```

## Description

Site level plots showing rainfall and reported flood events at each site. The dates extend 60 days in each direction (where possible) around the reported flood events.

In this example we set thresholds of `25 mm` on a `10 day` precipitation accumulation window. These are both arbitrary parameters, but were chosen for illustrative purposes to make sure the dates & events were being properly classified/labelled

- The event is labelled as a `TP` if rainfall crosses the threshold (25 mm) anytime between `7` days before and `3` days after the event was reported. 
- If the rainfall does not cross the threshold within the above time period it is a `FN`
- No consecutive dates greater than the threshold after a TP are labelled/classified
- For the rest of the dates greater than the threshold the first date of when the threshold is crossed is labelled as a `FP`. After the labelled FP, consecutive days above the threshold are note labelled.

**Note:** The window for plottin was set to 60 days around the reported flood events. Therefore, in many cases there are likely `FP`s that off the outside of the plot window.
 

```{r site-plots , echo=FALSE,out.width="100%"}
walk(p_all_sites_events_pred_classifications, ~ print(.x))
```

