---
title: "06_clustering events"
output: html_document
date: "2023-03-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include =T,warning=F,message = F)
```
## Clustering Events

```{r, echo=F,include=T}
library(here)
library(targets)
library(sf)
library(tidyverse)
library(leaflet)

library(gghdx)
library(lubridate )
library(sysfonts)
tar_source()

font_add_google("Source Sans Pro")
showtext::showtext_auto()

tar_load(cccm_flood_impact_data)

tar_load(cccm_flood_impact_data_w_coords)
tar_load(rainfall_impact_tbl)

coord_lookup <- cccm_flood_impact_data_w_coords %>% 
    distinct(site_id, lon, lat)

roi<- rainfall_impact_tbl %>% 
    left_join(coord_lookup, by ="site_id") %>% 
    filter(governorate_name %in% c("Marib","Hajjah"),fevent) 

unique_event_dates<- roi %>% 
    distinct(governorate_name,date) %>% 
    split(.$governorate_name)
```

Looking at just **Marib** and **Hajjah**. We see the following see that Marib and Hajjaah have **`r nrow(unique_event_dates$Marib)`** , and **`r nrow(unique_event_dates$Hajjah)`**, respectively.

Let's see if we can cluster based on coordinates and date of event. As the governorates are separated geographically let's do the clustering separately.

```{r,echo=F, include=T}
k <-  7
pts_clustered_k7 <- split(roi,roi$governorate_name) %>% 
    imap(\(df,nm){
        roi_matrix <- df %>% 
            mutate(
                date= as.numeric(date)
            ) %>% 
            select(date, lon, lat) %>% 
            as.matrix()
        km <-  kmeans(roi_matrix,centers= k,nstart = 20)
        df$cluster <- paste0(nm,"_",as.character(km$cluster))
        df_sf <-  st_as_sf(df,
                           coords = c("lon","lat"),
                           crs= 4326
        ) %>% 
            st_jitter(amount = 0.05)
        df_sf$labels <-  paste0(
            "Site_ID:", df_sf$site_id,
            "<br> date:",df_sf$date,
            "<br> cluter:", df_sf$cluster
        ) %>% lapply(htmltools::HTML)

        return(df_sf)
    }
    )

pal_1 <- colorFactor(palette = "Dark2", domain = pts_clustered_k7$Hajjah$cluster)
pal_2 <- colorFactor(palette = "Accent", domain = pts_clustered_k7$Marib$cluster)


k <-  5
pts_clustered_k5 <- split(roi,roi$governorate_name) %>% 
    imap(\(df,nm){
        roi_matrix <- df %>% 
            mutate(
                date= as.numeric(date)
            ) %>% 
            select(date, lon, lat) %>% 
            as.matrix()
        km <-  kmeans(roi_matrix,centers= k,nstart = 20)
        df$cluster <- paste0(nm,"_",as.character(km$cluster))
        df_sf <-  st_as_sf(df,
                           coords = c("lon","lat"),
                           crs= 4326
        ) %>% 
            st_jitter(amount = 0.05)
        df_sf$labels <-  paste0(
            "Site_ID:", df_sf$site_id,
            "<br> date:",df_sf$date,
            "<br> cluter:", df_sf$cluster
        ) %>% lapply(htmltools::HTML)
        
        grid <- st_make_grid(st_bbox(st_transform(df_sf,crs=2090)) %>% 
                         st_as_sfc(),
            cellsize = 5500) %>% 
            st_transform(crs= 4326)
        ret <- list()
        ret$grid <- grid
        ret$df_sf<-df_sf

        return(ret)
        
    }
    )

pal_1_k5 <- colorFactor(palette = "Dark2", domain = pts_clustered_k5$Hajjah$cluster)
pal_2_k5 <- colorFactor(palette = "Accent", domain = pts_clustered_k5$Marib$cluster)
```

I've jittered the points by 0.05 so that you it's easier to see the sites that are very close to one another

Here is a map with 7 clusters per governorate
```{r,echo=F,include=T,out.width="100%"}
leaflet() %>% 
    addTiles() %>% 
     addCircleMarkers(data= pts_clustered_k7$Hajjah,
                popup = ~cluster, 
                label = ~labels,
                # clusterOptions = markerClusterOptions(),
                color = ~pal_1(cluster)
             ) %>% 
     addCircleMarkers(data= pts_clustered_k7$Marib,
                popup = ~cluster, 
                label = ~labels,
                # clusterOptions = markerClusterOptions(),
                color = ~pal_2(cluster)
             )
```

and here is one with 5 clusters per governorate
```{r,echo=F, include =T,out.width="100%"}
?st_make_grid()
leaflet() %>% 
    addTiles() %>% 
    addProviderTiles(provider = leaflet::providers$Stamen.Toner) %>% 
    addPolygons(data=pts_clustered_k5$Hajjah$grid,fill=NA) %>%
    addPolygons(data=pts_clustered_k5$Marib$grid,fill=NA) %>%
     addCircleMarkers(data= pts_clustered_k5$Hajjah$df_sf,
                popup = ~cluster, 
                label = ~labels,
                # clusterOptions = markerClusterOptions(),
                color = ~pal_1_k5(cluster)
             ) %>% 
     addCircleMarkers(data= pts_clustered_k5$Marib$df_sf,
                popup = ~cluster, 
                label = ~labels,
                # clusterOptions = markerClusterOptions(),
                color = ~pal_2_k5(cluster)
             ) %>% 
    addScaleBar()
```

```{r}

pts_clustered_k5$Hajjah$df_sf %>% 
    st_drop_geometry( ) %>% 
    select(date, site_id,cluster) %>% 
    arrange(date) %>% 
    distinct(date,site_id,cluster) %>% 
    filter(site_id !="YE1712_0643") %>% 
    left_join(y = cccm_flood_impact_data %>%
                  group_by(site_id, date=date_of_episode) %>% 
                  summarise(
                      across(starts_with("num_"),~mean(.x,na.rm=T))
                  ) ,
              by = c("site_id"="site_id","date"="date")) %>% 
    print(n=nrow(.)) %>% 
    ggplot(aes(x=date,y=num_shelters_affected,color=cluster))+
    geom_point()+
    labs(y="Number affected Shelters",title = "Clustering of events based on timing and coordinates")+
    theme_hdx()

```