---
title: "09_ECMWF_MARS_"
output: 
    bookdown::html_document2:
        toc: true
        toc_float: true
        toc_depth: 4
date: "2023-04-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE, eval = TRUE
)
```


```{r libs}
library(tidyverse)
library(lubridate)
library(sf)
library(leaflet)
library(targets)
library(sysfonts)
library(gghdx)
library(here)
library(zoo)
library(extRemes)
library(data.table)
tar_source()

font_add_google("Source Sans Pro")
showtext::showtext_auto()

# ERA 5
tar_load(era5_w_rolling)

# MARS
tar_load(ecmwf_mars_leads_split_rolled)


# CHIRPS
tar_load(zonal_stats_high_risk_hull)
chirps <- zonal_stats_high_risk_hull %>% 
    map(~.x %>% 
            mutate(date=as_date(date)))
    

```


## Intro/background

In this doc:
- We look at ERA 5 historical precip data.
- Describe how the MARS data has been processed for forecast skill analysis
- calculate return periods for ERA5 historical precip data

## ERA5 Historical Rainfall.

Daily ERA5 historical precipitation was obtained from the `total_precipitation_sum` band of the [ERA5-Land Daily Aggregated - ECMWF Climate Reanalysis]("https://developers.google.com/earth-engine/datasets/catalog/ECMWF_ERA5_LAND_DAILY_RAW"). [ERA5 Daily Aggregates - Latest Climate Reanalysis Produced by ECMWF / Copernicus Climate Change Service]("https://developers.google.com/earth-engine/datasets/catalog/ECMWF_ERA5_DAILY") was also considered, but only had historical data up to `2020-07-09` so was not used.

ERA 5 historical precipitation data (1981-2022) were extracted to the area surrounding the high risk sites in Marib and Hajjah. The daily values in each were aggregated by weighted "mean" based on cell overlap with boundaries. Rolling 3, 5, and 10 day accumulation values were also calculated. The result is in the `era5_w_rolling` target. Displayed here:

```{r}
era5_w_rolling
```

## ECMWF Historical Forecast (MARS)

ECMWF MARS historical precipitation forecasts (2007-2022) were extracted to the area surrounding the high risk sites in Marib and Hajjah. The each raster originally contained 11 bands time steps: 0,1,2,3,4,5,6,7,8,9,10,11. The 0 time-step contained no data so was omitted. time steps 1:11 rep represent days 1-10. After extracting the weighted zonal means to the aoi polygons. The time-steps were converted to dates and split into columns in the target `ecmwf_mars_leads_split_rolled` displayed below. Additionally for each date-leadtime combination the 3,5, and 10 day rolling accumulation window calculations were made in addition to the daily values. **Note:** column `date_forecast_predict` in the target represents the date that is being forecasted, not the date the forecast was made.

```{r}

ecmwf_mars_leads_split_rolled
```


## ECMWF + ERA5

as you can see below the ERA5 & ECMWF MARS data can now easily be combined for skill analysis at different lead times as well as differen't accumulation windows
```{r}
mars_era5_combined <- ecmwf_mars_leads_split_rolled %>% 
    left_join(era5_w_rolling, 
              by =c("governorate_name","date_forecast_predict"="date")
              )
```

Here is single day accumulation with lead-time =1 
```{r}
mars_era5_combined %>% 
    ggplot(aes(x=lead_1_daily,y=era_precip_daily))+
    geom_point()+
    labs(x="ECMWF Forecast - Lead 1 (mm)",
         y="ERA5 Precipitation (mm)",
         title = "ECMWF Forecast vs ERA5 Historical",
         subtitle = "Daily Precipitation Values. Forecast set to a lead-time of 1 day"
         )+
    facet_wrap(~governorate_name)+
    theme_hdx()
```

We see that the correlation looks better when we aggregate to 3 day accumulation windows:
```{r}
mars_era5_combined %>% 
    ggplot(aes(x=lead_1_roll3,y=era_roll3))+
    geom_point()+
    facet_wrap(~governorate_name)+
    labs(x="ECMWF Forecast - Lead 1 (mm)",
         y="ERA5 Precipitation (mm)",
         title = "ECMWF Forecast vs ERA5 Historical",
         subtitle = "3 day rainfall acccumulation values. Forecast set to a lead-time of 1 day"
         )+
    theme_hdx()
```


```{r}

chirps_long <- chirps %>% 
    imap_dfr(\(df,nm){
        df %>% 
           mutate(
               name = paste0("chirps_",nm)
           ) %>% 
            rename(value = mean) %>% 
            select(-median)
    }
    )
chirps_era_long <- bind_rows(
    chirps_long,
    era5_w_rolling %>% 
    pivot_longer(cols = starts_with("era_"))
        
    
)
chirps_era_wide <- chirps_era_long %>% 
    pivot_wider(names_from=name, values_from = value)
```


## CHIRPS Vs ERA5 

### Rainfall scatter plots

```{r}
chirps_era_wide %>% 
    ggplot(aes(x= chirps_precip_daily,y=era_precip_daily))+
    geom_point(alpha=0.2)+
    facet_wrap(~governorate_name)+
    theme_hdx()+
    labs(
        x= "CHIRPS",
        y= "ERA5",
        title= "Daily Precip"
    )
chirps_era_wide %>% 
    ggplot(aes(x= chirps_roll3,y=era_roll3))+
    geom_point(alpha=0.2)+
    facet_wrap(~governorate_name)+
    theme_hdx()+
    labs(
        x= "CHIRPS",
        y= "ERA5",
        title= "3 day rain accumuation"
    )

chirps_era_wide %>% 
    ggplot(aes(x= chirps_roll5,y=era_roll5))+
    geom_point(alpha=0.2)+
    facet_wrap(~governorate_name)+
    theme_hdx()+
        labs(
        x= "CHIRPS",
        y= "ERA5",
        title= "5 day rain accumuation"
    )
chirps_era_wide %>% 
    ggplot(aes(x= chirps_roll10,y=era_roll10))+
    geom_point(alpha=0.2)+
    facet_wrap(~governorate_name)+
    theme_hdx()+
        labs(
        x= "CHIRPS",
        y= "ERA5",
        title= "10 day rain accumuation"
    )

```

## Timeseries Plots
```{r}
chirps_era_long %>%
    filter(str_detect(name,"_daily$")) %>% 
    ggplot(aes(x= date, y= value, color= name, group=name))+
    # ggplot(aes(x= chirps_roll10,y=era_roll10))+
    geom_line(alpha=0.5)+
    facet_wrap(~governorate_name)+
    theme_hdx()+
        labs(
        title= "Daily rain accumuation"
    )
chirps_era_long %>%
    filter(str_detect(name,"_roll3$")) %>% 
    ggplot(aes(x= date, y= value, color= name, group=name))+
    # ggplot(aes(x= chirps_roll10,y=era_roll10))+
    geom_line(alpha=0.5)+
    facet_wrap(~governorate_name)+
    theme_hdx()+
        labs(
        title= "3-day rain accumuation"
    )
chirps_era_long %>%
    filter(str_detect(name,"_roll5$")) %>% 
    ggplot(aes(x= date, y= value, color= name, group=name))+
    geom_line(alpha=0.5)+
    facet_wrap(~governorate_name)+
    theme_hdx()+
        labs(
        title= "5-day rain accumuation"
    )
chirps_era_long %>%
    filter(str_detect(name,"_roll10$")) %>% 
    ggplot(aes(x= date, y= value, color= name, group=name))+
    
    geom_line(alpha=0.5)+
    facet_wrap(~governorate_name)+
    theme_hdx()+
        labs(
        title= "10-day rain accumuation"
    )
```

## Return Periods


### ERA 5 Precip RPs

This chunk calculates the return period for the ERA data at daily, 3-day, 5-day, and 10-day accumulation levels.

```{r}
era_return_periods_df <- chirps_era_wide %>% 
    split(.$governorate_name) %>% 
    imap_dfr(
        \(df_temp,nm){
            c("era_precip_daily","era_roll3","era_roll5","era_roll10") %>% 
                map(\(precip_temp){
                    return_period_level_tibble(df = df_temp,
                           value = precip_temp,
                           date = "date",rp_year = c(2,3,4,5 )) %>% 
                        mutate(precip_regime= precip_temp,
                               governorate_name=nm)
                }
                ) 
            
        }
    ) 

era_return_periods_df
```

### ECMWF-MARS RPS

just the skeleton for ECMWF MARS RPs (not finished)

```{r}
tar_load(ecmwf_mars_leads_split_rolled)

ecmwf_mars_rps <- ecmwf_mars_leads_split_rolled %>% 
    split(.$governorate_name) %>% 
    imap_dfr(
        \(df_temp,nm){
            c("lead_1_daily","lead_8_daily") %>% 
                map(\(precip_temp){
                    return_period_level_tibble(df = df_temp,
                           value = precip_temp,
                           date = "date_forecast_predict",rp_year = c(2,3,4,5 )) %>% 
                        mutate(precip_regime= precip_temp,
                               governorate_name=nm)
                }
                ) 
            
        }
    ) 
```

## Compare GEE ERA5 Precip to Local

Just curious to make sure ERA5 precip from GEE is the same as our local rasters

```{r}
tar_load(era5_w_rolling)
# lets find the max rainfall date and look around that date as an example
era5_w_rolling %>% 
    filter(era_precip_daily==max(era_precip_daily))

# max precip occured in Sep 2001 - let's ppull the data from that month
era5_dir <-  file.path(Sys.getenv("AA_DATA_DIR"),"public", "raw","yem","ecmwf")
era5_fp <- file.path(era5_dir, "yem_era5_tp_2001-09.grib2")

# load month of data
era5_sep2001 <- terra::rast(era5_fp)

# rename layers with date
era5_times <- terra::time(era5_sep2001)
terra::set.names(era5_sep2001,terra::time(era5_sep2001))


# load convex hulls for extraction
tar_load(high_risk_hulls)

# extract data to hulls
era5_sep01_extr <- exact_extract(x = era5_sep2001,
              y=high_risk_hulls, 
              append_cols = "governorate_name",
    fun = c("mean"),
    force_df = T,
    full_colnames = T
  ) %>%
    pivot_longer(-matches("governorate_name")) %>%
    separate(name, into = c("stat", "date"), sep = "\\.") %>%
    pivot_wider(names_from = "stat", values_from = "value") %>% 
    mutate(
        mean = mean*1000 # m to mm
    )
 


# combine GEE version and local data and calculate difference in daily values
era5_daily_gee_vs_local <- era5_sep01_extr %>% 
    rename(
        date_time = "date"
    ) %>% 
    mutate(
        date= as_date(date_time)-1
    ) %>% 
    filter(str_detect(date_time,"00:00:00")) %>% 
    left_join(era5_w_rolling %>% select(governorate_name, era_precip_daily,date)) %>% 
    mutate(
        diff_mm=  mean- era_precip_daily
    )

# view 
era5_daily_gee_vs_local

# very little difference.
mean(era5_daily_gee_vs_local$diff_mm)

```

GEE ERA 5 daily and local raster are essentially the same. Differences likely due to slight differences in how weighted calculations/simplifications of polygons are performed -- could compare pixels, but don't think it is worth it at this point


