---
title: "09_ECMWF_MARS_"
output: 
    bookdown::html_document2:
        toc: true
        toc_float: true
        toc_depth: 4
date: "2023-04-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE, eval = TRUE
)
```


```{r libs}
library(tidyverse)
library(lubridate)
library(glue)
library(targets)
library(gghdx)
library(sysfonts)           # req for gghdx
library(here)
library(zoo)
library(extRemes)           # RP calcs
library(data.table)
library(ggridges)           # ridge plot
library(ggiraph)            # interactive ggplot 
library(exactextractr)      # fast weighted zonal stats
library(sf)                 # needed on backend for clustering points in last plots (impact w Rps)
library(terra)              # raster
tar_source() # source funcs

font_add_google("Source Sans Pro")
showtext::showtext_auto()

# ERA 5
tar_load(era5_w_rolling) # from GEE
tar_load(era5_rolling_zonal_local) # from local files - `00:00` band

# MARS
tar_load(ecmwf_mars_leads_split_rolled)

# CHIRPS
tar_load(zonal_stats_high_risk_hull)
chirps <- zonal_stats_high_risk_hull %>% 
    map(~.x %>% 
            mutate(date=as_date(date)))


tar_load(era_chirps_long) # for plotting chirps era together
    

```


## Intro/background

In this doc
- Describe the following targets
- We look at `ECMWF ERA 5` historical precipitation data
- We look at `ECMWF HRES (MARS)` forecast/hindcast precipitaiton data
- Preliminary attempt to look at accuracy of forecast per lead time
- calculate return periods for ERA5 historical precip data

### ERA5 Historical Rainfall.

Daily ERA5 historical precipitation was obtained from the `total_precipitation_sum` band of the [ERA5-Land Daily Aggregated - ECMWF Climate Reanalysis]("https://developers.google.com/earth-engine/datasets/catalog/ECMWF_ERA5_LAND_DAILY_RAW"). [ERA5 Daily Aggregates - Latest Climate Reanalysis Produced by ECMWF / Copernicus Climate Change Service]("https://developers.google.com/earth-engine/datasets/catalog/ECMWF_ERA5_DAILY") was also considered, but only had historical data up to `2020-07-09` so was not used.

ERA 5 historical precipitation data (1981-2022) were extracted to the area surrounding the high risk sites in Marib and Hajjah. The daily values in each were aggregated by weighted "mean" based on cell overlap with boundaries. Rolling 3, 5, and 10 day accumulation values were also calculated. The result is in the `era5_w_rolling` target. We a;sp performed weighted mean stats using local ERA rasters stored on the drive. Here we merge them to show tht the results are the same - minimal variations could be due to a variety of factors including how the weighted stats were calculated. 

```{r}
era5_w_rolling %>% 
    left_join(era5_rolling_zonal_local %>% 
                  mutate(
                      date= as_date(date),
                      across(matches("^roll|^precip_"),~.x*1000)
                      )
              ) %>% 
    select(governorate_name, date, GEE_ERA = era_precip_daily,local_ERA = precip_daily)
```

### ECMWF HRES Forecast/Hindcast (MARS)

Initially I was curious in trying to assess if any particular lead times were more accurate than others just to learn a little more about the forecast data. Therefore I tried splitting the data by lead time and calculating rolling statistics on predicted forecast per lead time. This was done to create: `ecmwf_mars_leads_split_rolled` - explained more below:

ECMWF MARS historical precipitation forecasts (2007-2022) were extracted to the area surrounding the high risk sites in Marib and Hajjah. The each raster originally contained 11 bands time steps: 0,1,2,3,4,5,6,7,8,9,10,11. The 0 time-step contained no data so was omitted. time steps 1:11 rep represent days 1-10. After extracting the weighted zonal means to the aoi polygons. The time-steps were converted to dates and split into columns in the target `ecmwf_mars_leads_split_rolled` displayed below. Additionally for each date-leadtime combination the 3,5, and 10 day rolling accumulation window calculations were made in addition to the daily values. **Note:** column `date_forecast_predict` in the target represents the date that is being forecasted, not the date the forecast was made.

```{r}
ecmwf_mars_leads_split_rolled
```

I also generated a forecast rolling statistics data.frame more in line with how the forecast would also be triggered in target: `hres_rolled_per_date_gen`. In this target forecasted precip is rolled per date forecast made/generated. This is the target that makes more sense to use for trigger simulation/assessment. I made this later after writing most of this document so will save more detailed discussion of analysis on this for another document.

### ECMWF + ERA5

as you can see below the ERA5 & ECMWF MARS data can now easily be combined for skill analysis at different lead times as well as different accumulation windows

```{r}
mars_era5_combined <- ecmwf_mars_leads_split_rolled %>% 
    left_join(era5_w_rolling, 
              by =c("governorate_name","date_forecast_predict"="date")
              )
```

Here is single day accumulation with lead-time =1 
```{r}
mars_era5_combined %>% 
    ggplot(aes(x=lead_1_daily,y=era_precip_daily))+
    geom_point()+
    labs(x="ECMWF Forecast - Lead 1 (mm)",
         y="ERA5 Precipitation (mm)",
         title = "ECMWF Forecast vs ERA5 Historical",
         subtitle = "Daily Precipitation Values. Forecast set to a lead-time of 1 day"
         )+
    facet_wrap(~governorate_name)+
    theme_hdx()
```

We see that the correlation looks better when we aggregate to 3 day accumulation windows:
```{r}
mars_era5_combined %>% 
    ggplot(aes(x=lead_1_roll3,y=era_roll3))+
    geom_point()+
    facet_wrap(~governorate_name)+
    labs(x="ECMWF Forecast(mm) - Lead time = 1 day" ,
         y="ERA5 Precipitation (mm)",
         title = "ECMWF - HRES (Forecast) vs ECMWSF - ERA5 (Historical)",
         subtitle = "3 day rainfall acccumulation values. Forecast set to a lead-time of 1 day"
         )+
    theme_hdx()
```


## ECMWF HRES by lead time.

Below we look at predicted rainfall as a function of lead time for both 1 day and 3 day accumulations in fig \@ref(fig:plot-HRESdailyRidge) and \@ref(fig:plot-HRES3dRidge), respectively. In the 3 day rainfall plot the values are roller per date genreated to be in line with how values would be rolled for trigger. 

```{r plot-HRESdailyRidge, fig.cap="Forecasted daily rainfall (mm) as function of lead time (days). Values < 1 mm removed for visualization, pseudolog scaling applied to x axis",results = "asis"}
# rm < 1 - looks a bit noisy!
tar_load(hres_rolled_per_date_gen)
hres_rolled_per_date_gen %>% 
    filter(value>1) %>% 
    filter(str_detect(name,"precip_daily")) %>% 
    ggplot(aes(x= value,y= as_factor(leadtime)))+
    geom_density_ridges()+
    scale_x_continuous(breaks = c(0,2,4,6,8,10,20,30,40,50,100,200,500,1000,1500),
                       trans = scales::pseudo_log_trans()
                       )+
    labs(x="Forecasted Daily Rainfall (mm)",
         y= "lead time (days)",
         title = "Forecasted rainfall as function of lead time")+
    theme_hdx()+
     theme(
        axis.text.x = element_text(angle = 90)
    )+
    facet_wrap(~governorate_name)
```

```{r plot-HRES3dRidge, fig.cap="Forecasted 3 day rainfall accumulationn (mm) as function of lead time (days). Values < 1 mm removed for visualization, pseudolog scaling applied to x axis",results = "asis"}
hres_rolled_per_date_gen %>% 
    filter(value>1) %>% 
    filter(str_detect(name,"roll3")) %>% 
    ggplot(aes(x= value,y= as_factor(leadtime)))+
    geom_density_ridges()+
    scale_x_continuous(breaks = c(0,1,2,4,6,8,10,15,20,30,40,50,100,200,500,1000,1500),
                       trans = scales::pseudo_log_trans()
                       )+
    labs(x="Forecasted 3-Day Rainfall (mm)",
         y= "lead time (days)",
         title = "Forecasted rainfall as function of lead time")+
    theme_hdx()+
    theme(
        axis.text.x = element_text(angle = 90)
    )+ 
    facet_wrap(~governorate_name)



```


## CHIRPS Vs ERA5 

### Timeseries Plots
```{r}
era_chirps_long %>%
    filter(str_detect(precip_regime,"_daily$")) %>% 
    ggplot(aes(x= date, y= value, color= source, group=source))+
    # ggplot(aes(x= chirps_roll10,y=era_roll10))+
    geom_line(alpha=0.5)+
    facet_wrap(~governorate_name)+
    theme_hdx()+
        labs(
        title= "Daily rain accumuation"
    )


era_chirps_long %>%
    filter(str_detect(precip_regime,"roll3$")) %>% 
    ggplot(aes(x= date, y= value, color= source, group=source))+
    geom_line(alpha=0.5)+
    facet_wrap(~governorate_name)+
    theme_hdx()+
        labs(
        title= "3-day rain accumuation"
    )

era_chirps_long %>%
    filter(str_detect(precip_regime,"roll5$")) %>% 
    ggplot(aes(x= date, y= value, color= source, group=source))+
    geom_line(alpha=0.5)+
    facet_wrap(~governorate_name)+
    theme_hdx()+
        labs(
        title= "5-day rain accumuation"
    )

era_chirps_long %>%
    filter(str_detect(precip_regime,"roll10$")) %>% 
    ggplot(aes(x= date, y= value, color= source, group=source))+
    
    geom_line(alpha=0.5)+
    facet_wrap(~governorate_name)+
    theme_hdx()+
        labs(
        title= "10-day rain accumuation"
    )
```

### Rainfall scatter plots

I want to plot ERA vs CHIR
None of these are particularly compelling?

```{r}

p_marib_era_vs_chirps_pct_diff <- era_chirps_long %>% 
    filter(!is.nan(value)) %>% 
    mutate(value = value +1) %>% 
    group_by(source,governorate_name,precip_regime) %>% 
    arrange(date) %>% 
    mutate(
        diff= value-lag(value),
        pct_diff= diff/lag(value)
    ) %>% 
    select(-diff,-value) %>% 
    filter(governorate_name=="Marib") %>% 
    pivot_wider(values_from = "pct_diff",names_from="source") %>% 
    ggplot(aes(x= ERA,y=CHIRPS))+
    geom_point()+
    facet_wrap(~precip_regime)+
    labs(title= "ERA vs Chirps",subtitle = "Marib: pct difference")+
    theme_hdx()


p_hajjah_era_vs_chirps_pct_diff <- era_chirps_long %>% 
    filter(!is.nan(value)) %>% 
    mutate(value = value +1) %>% 
    group_by(source,governorate_name,precip_regime) %>% 
    arrange(date) %>% 
    mutate(
        diff= value-lag(value),
        pct_diff= diff/lag(value)
    ) %>% 
    select(-diff,-value) %>% 
    filter(governorate_name=="Hajjah") %>% 
    pivot_wider(values_from = "pct_diff",names_from="source") %>% 
    ggplot(aes(x= ERA,y=CHIRPS))+
    geom_point()+
    facet_wrap(~precip_regime)+
    labs(title= "ERA vs Chirps",subtitle = "Hajjah")+
    theme_hdx()


p_marib_era_vs_chirps_pct_diff

p_hajjah_era_vs_chirps_pct_diff

```



# ERA5 - Flood Impact & RPs

```{r}
# some quick prep for the `plot_rainfall_rps_impact` function which is not super generalizeable
tar_load(cccm_flood_marib_hajjah_impact)
era5_long_split <- era5_w_rolling %>% 
    pivot_longer(cols = matches(".*roll*|.*daily"),values_to = "mean") %>% 
    mutate(name = str_remove(name,"era_")) %>% 
    split(.$name)


p_era5_rps_impact <- era5_long_split %>% 
    names() %>% 
    map(\(precip_windows){
        plot_rainfall_rps_impact(impact_data=cccm_flood_marib_hajjah_impact,
                                 historical_rainfall=era5_long_split,
                                 precip_regime= precip_windows, # map through each precip regime
                                 impact_var = "num_shelters_affected",
                                 rp_year= c(2,3,4,5, 10), # return period years
                                 scale=F, # don't scale vars before kmeans
                                 k=5, # 5 clusters 
                                 aggregate_impact = NULL # plot individual event impact -not aggregated
        )
        
    }) %>% 
    set_names(era5_long_split %>%
                  names() )

p_era5_rps_impact$roll3$Hajjah
p_era5_rps_impact$roll3$Marib
```


## Appendix

### Compare GEE ERA5 Precip to Local

Just curious to make sure ERA5 precip from GEE is the same as our local rasters

```{r}
tar_load(era5_w_rolling)
# lets find the max rainfall date and look around that date as an example
era5_w_rolling %>% 
    filter(era_precip_daily==max(era_precip_daily))

# max precip occured in Sep 2001 - let's ppull the data from that month
era5_dir <-  file.path(Sys.getenv("AA_DATA_DIR"),"public", "raw","yem","ecmwf")
era5_fp <- file.path(era5_dir, "yem_era5_tp_2001-09.grib2")

# load month of data
era5_sep2001 <- terra::rast(era5_fp)

# rename layers with date
era5_times <- terra::time(era5_sep2001)
terra::set.names(era5_sep2001,terra::time(era5_sep2001))


# load convex hulls for extraction
tar_load(high_risk_hulls)

# extract data to hulls
era5_sep01_extr <- exact_extract(x = era5_sep2001,
              y=high_risk_hulls, 
              append_cols = "governorate_name",
    fun = c("mean"),
    force_df = T,
    full_colnames = T
  ) %>%
    pivot_longer(-matches("governorate_name")) %>%
    separate(name, into = c("stat", "date"), sep = "\\.") %>%
    pivot_wider(names_from = "stat", values_from = "value") %>% 
    mutate(
        mean = mean*1000 # m to mm
    )
 


# combine GEE version and local data and calculate difference in daily values
era5_daily_gee_vs_local <- era5_sep01_extr %>% 
    rename(
        date_time = "date"
    ) %>% 
    mutate(
        date= as_date(date_time)-1
    ) %>% 
    filter(str_detect(date_time,"00:00:00")) %>% 
    left_join(era5_w_rolling %>% select(governorate_name, era_precip_daily,date)) %>% 
    mutate(
        diff_mm=  mean- era_precip_daily
    )

# view 
era5_daily_gee_vs_local

# very little difference.
mean(era5_daily_gee_vs_local$diff_mm)

```

GEE ERA 5 daily and local raster are essentially the same. Differences likely due to slight differences in how weighted calculations/simplifications of polygons are performed -- could compare pixels, but don't think it is worth it at this point


### ECMWF HRES - Negative pixels
WHY SO MANY negative values in the ecmwf?

```{r}

ecmwf_mars_dir <-  file.path(
    Sys.getenv("AA_DATA_DIR"),
    "private",
    "processed",
    "yem",
    "ecmwf",
    "strange_coords"
)


r_rand<- terra::rast(list.files(ecmwf_mars_dir,full.names = T)[2])
r_rand_sub <- r_rand["tp_step=24$"]
hist(r_rand_sub)
neg_value_counts <- data.frame(r_vals=terra::values(r_rand_sub) %>% 
    as.numeric()
) %>% 
    filter(r_vals<0) %>% 
    count(r_vals) # some weird R numeric float stuff?

# some weird R float precision stuff?
neg_value_counts$r_vals[1]-neg_value_counts$r_vals[2]


```

### Order of Operations Experiment

if we have daily rasters are the following equivalent:

a.) run a zonal mean (reducer) on all images and then roll those point values (i.e rolling sum)
b.) roll raster pixels and then run zonal mean on rolled rasters.


It appears it does not matter!

below we compare:

- `era5_rolling_zonal_local` - pixels rolled and then zonal mean reducer performed on each (data source: local raster)
- `era5_w_rolling` - zonal mean performed on daily, mean point values rolled

```{r}

era5_rolling_zonal_local %>% 
    mutate(
        date=as_date(date),
        across(where(is.numeric),~.x*1000)   
           )%>% 
    left_join(
        era5_w_rolling,
        by =c("governorate_name","date")) %>% 
    select(date, contains("precip"),matches(".*roll3$")) %>% 
    print(n=100)
```


