---
title: "3 District Investigation"
subtitle: "Abs, Ma'rib, & Ma'rib City"
output: 
    bookdown::html_document2:
        toc: true
        toc_float: true
        toc_depth: 4
date: "2023-05-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  eval = TRUE,
  results = "asis"
)
```

```{r libs}
library(tidyverse)
library(lubridate)
library(sf)
library(targets)
library(leaflet)
library(extRemes)
library(gt)
library(gghdx)
gghdx()
tar_source()

# tar_load(cccm_flood_marib_hajjah_impact) # impa
tar_load(adm_sf) # admin units (list of sf features)
tar_load(high_risk_hulls) # high risk hulls calculated in targets from high risk sites in 2 govs
tar_load(cccm_wb) # cccm master list of sites
tar_load(cccm_floodscore_df) # list of sites with flood scores
```

## Take-away Summary:

Adjusting trigger thresholds and AOI for forecast monitoring to just high risk sites in 3 districts (within Hajjah & Ma'rib governorate) appears justified. Here is the rationale:

- Trigger development done for all high risk sites in the 2 Governorates (Hajjah, Ma'rib) so that we would have sufficient historical impact data to evaluate trigger thresholds
- We found a trigger level that was acceptable (2 year 3day cumulative RP event )
- We adjusted the AOI to a smaller geographic subset (3 districts) based on where partners will respond and assume the same relation b/w impact and rainfall
- We calculate new 2 year RP level for the new AOI and take the lowest values of the 2 analyses to be conservative.         + New rounded thresholds: **Hajjah: 24 mm and Ma'rib: 20 mm**

## Intro

This script explores adjusting the AOI and how that would impact threhsolds. We do some exploration which was used to build various targets. Those targets are then imported and explored.

## Background

Initially did analysis based on high risk sites in Hajjah and Ma'rib. By using all high risk sites we were able to compare historical rainfall to events. Later the priority site list was filtered to just sites in 3 districts inside Hajjah and Ma'rib: Abs (Hajjah), Ma'rib, & Ma'rib City (both Ma'rib). We had initially decided to continue with the analysis based on the all high risk sites with Hajjah and Ma'rib governorates (convex hull of all) for multiple reasons:

-list of priority sites contained
    + sites not in master list
    + sites without coordinate
    + sites that were not classified as high risk
- If we filter sites with coordinates in new priority list we loose much of the ability to compare with historical data because the number of sites is greatly reduced

Nonetheless, **purpose** of this document is just to compare 2 year RP that we would get if we included only high-risk sites in Hajjah & Ma'rib. We may want to consider changing thresholds based on this.

```{r}
#########################################################################
# **Note**
# the below code has been re-factored and added to the targets pipeline in
# func: `extract_sites_w_coords_by_floodscore()`
# target: `high_risk_site_pts`
#########################################################################


# clean up adm2 - remove unnecessary cols
adm2 <- adm_sf$yem_admbnda_adm2_govyem_cso_20191002 %>%
  select(matches("^adm\\d_[ep]"))


# only master list sites with coords
master_db_coords <- cccm_wb$`ML- Flooding Available data` %>%
  select(
    longitude = available_coordinates_longitude,
    latitude = available_coordinates_latitude, site_id
  ) %>%
  filter(!is.na(longitude), !is.na(latitude))

# only high risk sites w/ coords
high_risk_db_sp <- cccm_floodscore_df %>%
  filter(site_flood_hazard_score == "High Hazard") %>%
  left_join(master_db_coords) %>%
  filter(!is.na(longitude), !is.na(latitude)) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326)

# join to clean admin boundaries
high_risk_db_sp <- high_risk_db_sp %>%
  st_join(adm2)


# figs for technical note n=94 & n=56
# high_risk_db_sp %>% 
#     filter(governorate_name %in% c("Marib","Hajjah"))
#     filter(adm2_en %in% c("Ma'rib","Ma'rib City","Abs")) %>% 
#     # nrow()
#     count(adm2_en)
    
```
 

 
 
````{r} 
# - `high_risk_hulls` - made from all high risk sites with `governorate_name` of `Hajjah` or `Marib` except site_id `YE1712_0643` which was clearly outside
# - `gov_hulls` made from all high risk sites falling in the 2 governorates after spatial joining with the admin
# - `district_hull` made from all high risk sies falling the the 3 districts (Abs, Marib, Marib City) based on spatial join with admin

#############################################################################
# **NOTE** this is dev code and has been refactored and added into the targets pipeline in
# target: `high_risk_district_hulls`
# func: `gen_district_hulls()`
# you will see tar_load(high_risk_district_hulls)  - in chunk below
#############################################################################


gov_hulls <- c("Hajjah", "Ma'rib") %>%
  map(
    \(gov){
      pts_sf <- high_risk_db_sp %>%
        filter(
          adm1_en == gov,
        ) %>%
        st_as_sf(coords = c("longitude", "latitude"), crs = 4326)
      st_convex_hull(st_union(pts_sf)) %>%
        st_as_sf() %>%
        mutate(governorate_name = gov) %>%
        rename(geometry = x)
    }
  ) %>%
  bind_rows()

district_hulls <- c("Abs", "Ma'rib") %>%
  map(
    \(district){
      pts_sf <- high_risk_db_sp %>%
        filter(
          str_detect(adm2_en, district) # this will catch Ma'rib and Ma'rib City
        ) %>%
        st_as_sf(coords = c("longitude", "latitude"), crs = 4326)
      st_convex_hull(st_union(pts_sf)) %>%
        st_as_sf() %>%
        mutate(dist = district) %>%
        rename(geometry = x)
    }
  ) %>%
  bind_rows()
```



## Current AOI vs AOI defined from 3 Districts
-Let's compare the difference:  <span style="color: orange;">orange - from 3 districts</span>,<span style="color: green;">previous aoi </span>
```{r}
leaflet() %>%
    # addTiles() %>% 
    addProviderTiles(provider = providers$Esri.WorldShadedRelief) %>% 
    addPolygons(data = adm_sf$yem_admbnda_adm1_govyem_cso_20191002, fill = NA, fillColor = NA, color="black") %>%
    # addPolygons(data = adm_sf$yem_admbnda_adm2_govyem_cso_20191002, color = "grey", fill = NA, fillColor = NA) %>%
    addPolygons(data = high_risk_hulls, color = "green", fillColor = NA, fillOpacity = 0) %>% 
    addPolygons(data = district_hulls, color = "orange", fillColor = "orange", fillOpacity = 1) 

```


```{r eval=F}
#####################################
# this chunk was a bit of an after thought it is just to check the approximate number of chirps pixels
# for both AOI's. We won't run it in the knitted version - eval=F
#######################################

library(rgee)
ee_Initialize()
chirps_link <- "UCSB-CHG/CHIRPS/DAILY"
chirps <- ee$ImageCollection(chirps_link)
july_rain <- as_tidyee(chirps$filterDate("2021-07-01", "2021-07-22"))

july_sum_ee_img <- chirps$filterDate("2021-07-01", "2021-07-22")$sum()
july_sum <- july_rain %>%
  # group_by(year, month) %>%
  summarise(
    stat = "sum"
  )

dist_hull_ee <- sf_as_ee(high_risk_district_hulls)
gov_hull_ee <- sf_as_ee(high_risk_hulls)

countDict_dist <- ee$Image$reduceRegions(
  image = july_sum_ee_img,
  reducer = ee$Reducer$count(),
  collection = dist_hull_ee,
  scale = 5500
)
countDict_gov <- ee$Image$reduceRegions(
  image = july_sum_ee_img,
  reducer = ee$Reducer$count(),
  collection = gov_hull_ee,
  scale = 5500
)
countDict_dist$getInfo()
countDict_gov$getInfo()
```

## Note: Current AOI vs AOI defined from Shapefile Boundaries

- It's worth quickly noting that I originally created the gov level AOI using the `governorate_name` in the master list from CCCM. This may or may not have been a good decision. On the one hand it could be influenced by typos, on the other hand gov names associated with sites by CCCM might be more up to date and/or more operationally relevant than the admin boundaries from shapefile
- Let's compare the difference:  <span style="color: orange;">orange - from shapefile</span>,<span style="color: green;">green - from master list name</span>
-slightly different, but nothing inherently wrong

```{r}
tar_load(high_risk_district_hulls)
leaflet() %>%
  addTiles() %>%
  addPolygons(data = adm_sf$yem_admbnda_adm1_govyem_cso_20191002, fill = NA, fillColor = NA) %>%
  addPolygons(data = adm_sf$yem_admbnda_adm2_govyem_cso_20191002, color = "grey", fill = NA, fillColor = NA) %>%
  addPolygons(data = gov_hulls, color = "orange", fillColor = NA, fillOpacity = 0) %>%
  addPolygons(data = high_risk_hulls, color = "green")
```

## CHIRPS Comparison

```{r}
#######################
# **note** `zonal_stats_high_risk_district_hull` created in targets
# and is zonal stats for CHIRPS with new AOI
#######################

tar_load(zonal_stats_high_risk_district_hull)

# calculate RP levels for new AOI
new_aoi_chirps_rps_full_record <- zonal_stats_high_risk_district_hull$roll3 %>%
  split(.$governorate_name) %>%
  imap_dfr(\(df_temp, nm){
    # cat(nm,"/n")
    return_period_level_tibble(
      df = df_temp,
      date = "date",
      value = "mean",
      rp_year = c(2, 3, 4, 5)
    ) %>%
      mutate(governorate_name = nm)
  })

tar_load(zonal_stats_high_risk_hull)

# RP levels for old AOI
old_aoi_chirps_rps_full_record <- zonal_stats_high_risk_hull$roll3 %>%
  split(.$governorate_name) %>%
  imap_dfr(\(df_temp, nm){
    # cat(nm,"/n")
    return_period_level_tibble(
      df = df_temp,
      date = "date",
      value = "mean",
      rp_year = c(2, 3, 4, 5)
    ) %>%
      mutate(governorate_name = nm)
  })
```


```{r, results="asis"}
# table of old AOI RP level 2
old_aoi_chirps_rps_full_record %>%
  filter(RP == 2) %>%
  gt() %>%
  fmt_number(
    columns = everything(),
    decimals = 1
  ) %>%
  fmt_number(
    columns = RP,
    decimals = 0
  ) %>%
  tab_header(
    title = "2 year RP level CHIRPS (1988-2022) by governorate",
    subtitle = "All high risk sites in Hajjah and Ma'rib governorates"
  ) %>%
  opt_align_table_header(align = c("left", "center", "right"))
```

```{r, results="asis"}
# table of new AOI RP level 2
new_aoi_chirps_rps_full_record %>%
  filter(RP == 2) %>%
  gt() %>%
  fmt_number(
    columns = everything(),
    decimals = 1
  ) %>%
  fmt_number(
    columns = RP,
    decimals = 0
  ) %>%
  tab_header(
    title = "2 year RP level CHIRPS (1988-2022) by governorate",
    subtitle = "All high risk sites in Abs (Hajjah) and Ma'rib/Ma'rib City (Ma'rib)"
  ) %>%
  opt_align_table_header(align = c("left", "center", "right"))
```


```{r}
# targets created using chirps-gefs and new aoi (zonal stast)
tar_load(gefs_zonal_districts_rolled)
tar_load(gefs_zonal_rolled)


new_gefs_rps <- gefs_zonal_districts_rolled %>%
  filter(name == "roll3") %>%
  filter(
    !leadtime %in% c(1:2), # rm lt 1-2 since they don't exist in 3-day rolling sum
    year(date_forecast_predict) != 2023, # rm 2023 since not complete
    year(date_forecast_predict) != 2020, # rm 2023 since incomplete data
  ) %>%
  mutate(
    gov_lt = paste0(governorate_name, "_", leadtime)
  ) %>%
  # calculate RP levels for each gov separately
  split(.$gov_lt) %>%
  imap_dfr(\(df_temp, nm){
    # cat(nm,"/n")
    return_period_level_tibble(
      df = df_temp,
      date = "date_forecast_predict",
      value = "value",
      rp_year = c(2, 3, 4, 5)
    ) %>%
      mutate(gov_lt = nm) %>%
      separate(col = gov_lt, into = c("governorate_name", "leadtime"))
  })

# same as above - but for old AOI
old_gefs_rps <- gefs_zonal_rolled %>%
  filter(name == "roll3") %>%
  filter(
    !leadtime %in% c(1:2),
    year(date_forecast_predict) != 2023,
    year(date_forecast_predict) != 2020,
  ) %>%
  mutate(
    gov_lt = paste0(governorate_name, "_", leadtime)
  ) %>%
  split(.$gov_lt) %>%
  imap_dfr(\(df_temp, nm){
    # cat(nm,"/n")
    return_period_level_tibble(
      df = df_temp,
      date = "date_forecast_predict",
      value = "value",
      rp_year = c(2, 3, 4, 5)
    ) %>%
      mutate(gov_lt = nm) %>%
      separate(col = gov_lt, into = c("governorate_name", "leadtime"))
  })
```


```{r}
# we already calculate CHIRPS historical RPs for 1988-2022
# but let's filter it first in the same manner as CHIRPS-GEFS so we can
# better compare the 2

# get chirps RP for each area (using new AOI)
new_chirps_2000_2022_rp <- zonal_stats_high_risk_district_hull$roll3 %>%
  filter(year(date) %in% c(2000:2022)) %>%
  filter(year(date) != 2021) %>%
  split(.$governorate_name) %>%
  imap_dfr(\(df_temp, nm){
    # cat(nm,"/n")
    return_period_level_tibble(
      df = df_temp,
      date = "date",
      value = "mean",
      rp_year = c(2, 3, 4, 5)
    ) %>%
      mutate(governorate_name = nm)
  })

# same as above using old AOI
old_chirps_2000_2022_rp <- zonal_stats_high_risk_hull$roll3 %>%
  filter(year(date) %in% c(2000:2022)) %>%
  filter(year(date) != 2021) %>%
  split(.$governorate_name) %>%
  imap_dfr(\(df_temp, nm){
    # cat(nm,"/n")
    return_period_level_tibble(
      df = df_temp,
      date = "date",
      value = "mean",
      rp_year = c(2, 3, 4, 5)
    ) %>%
      mutate(governorate_name = nm)
  })
```


```{r, results="asis"}
# table of new AOI  - title gives more info
# filter to just 2 year return period
new_chirps_2000_2022_rp %>%
  filter(RP == 2) %>%
  gt() %>%
  fmt_number(
    columns = everything(),
    decimals = 1
  ) %>%
  fmt_number(
    columns = RP,
    decimals = 0
  ) %>%
  tab_header(
    title = "2 year RP level CHIRPS (2000-2022) by governorate",
    subtitle = "All high risk sites in Abs (Hajjah) and Ma'rib/Ma'rib City (Ma'rib)"
  ) %>%
  opt_align_table_header(align = c("left", "center", "right"))
```


```{r,results="asis"}
# table of old AOI  - title gives more info
old_chirps_2000_2022_rp %>%
  filter(RP == 2) %>%
  gt() %>%
  fmt_number(
    columns = everything(),
    decimals = 1
  ) %>%
  fmt_number(
    columns = RP,
    decimals = 0
  ) %>%
  tab_header(
    title = "2 year RP level CHIRPS (2000-2022) by governorate",
    subtitle = "All high risk sites in Hajjah & Ma'rib governorates"
  ) %>%
  opt_align_table_header(align = c("left", "center", "right"))
```

Based on the table below we can arrive at new thresholds. Take the rounded lower value between the old and new AOI:
- Hajjah: 24
- Ma'rib: 20

```{r, results="asis"}
# calc mean RP 2 for LT 3-7. This is the activation window

# for old AOI
old_gefs_mean_rp <- old_gefs_rps %>%
  filter(
    RP == 2,
    leadtime %in% c(3:7)
  ) %>%
  group_by(governorate_name) %>%
  summarise(
    `previous 2 year RP` = mean(estimate)
  )

# same but for new AOI
new_gefs_mean_rp <- new_gefs_rps %>%
  filter(
    RP == 2,
    leadtime %in% c(3:7)
  ) %>%
  group_by(governorate_name) %>%
  summarise(
    `new 2 year RP` = mean(estimate)
  )

# tab_header title explains it all
left_join(old_gefs_mean_rp, new_gefs_mean_rp) %>%
  gt() %>%
  fmt_number(
    columns = everything(),
    decimals = 1
  ) %>%
  tab_header(
    title = "2 year RP level CHIRPS-GEFS (2000-2022) by governorate",
    subtitle = "Old method (all gov) vs new method (3 districts)"
  ) %>%
  opt_align_table_header(align = c("left", "center", "right"))
```

## Note - Targets Added

analysis above was added as targets so they are easily accessible:

1. a target for convex hull of Abs high risk sites:
    func:   `gen_district_hulls()` 
    target:`high_risk_district_hulls`
2. a target for chirps extraction of high-risk sites. 
    target chirps: `zonal_stats_high_risk_district_hull`
3. a target for chirps-gefs extraction.
    target chirps-gefs : `gefs_zonal_districts_rolled`
