---
title: "3 District Investigation"
subtitle: "Abs, Ma'rib, & Ma'rib City"
output: 
    bookdown::html_document2:
        toc: true
        toc_float: true
        toc_depth: 4
date: "2023-05-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  eval = TRUE,
  results="asis"
)
```

```{r libs}
library(tidyverse)
library(lubridate)
library(sf)
library(targets)
library(leaflet)
library(extRemes)
library(gt)
library(gghdx)
gghdx()
tar_source()

tar_load(cccm_flood_marib_hajjah_impact)
tar_load(adm_sf)
tar_load(high_risk_hulls)
tar_load(cccm_wb)
tar_load(cccm_floodscore_df)
```

## Background

Initially did analysis based on high risk sites in Hajjah and Ma'rib. By using all high risk sites we were able to compare historical rainfall to events. Later the priority site list was filtered to just sites in 3 districts inside Hajjah and Ma'rib: Abs (Hajjah), Ma'rib, & Ma'rib City (both Ma'rib). We had initially decided to continue with the analysis based on the all high risk sites with Hajjah and Ma'rib governorates (convex hull of all) for multiple reasons:

-list of priority sites contained
    + sites not in master list
    + sites without coordinate
    + sites that were not classified as high risk
- If we filter sites with coordinates in new priority list we loose much of the ability to compare with historical data because the number of sites is greatly reduced

Nonetheless, **purpose** of this document is just to compare 2 year RP that we would get if we included only high-risk sites in Hajjah & Ma'rib. We may want to consider changing thresholds based on this.

```{r}

adm2 <- adm_sf$yem_admbnda_adm2_govyem_cso_20191002 %>% 
    select(matches("^adm\\d_[ep]"))


master_db_coords <- cccm_wb$`ML- Flooding Available data` %>% 
    select(
      longitude = available_coordinates_longitude,
      latitude = available_coordinates_latitude, site_id
    ) %>%
    filter(!is.na(longitude), !is.na(latitude))

high_risk_db_sp <- cccm_floodscore_df %>% 
    filter(site_flood_hazard_score == "High Hazard") %>%
    left_join(master_db_coords) %>%
    filter(!is.na(longitude), !is.na(latitude)) %>% 
    st_as_sf(coords=c("longitude","latitude"),crs=4326)

high_risk_db_sp <- high_risk_db_sp %>% 
    st_join(adm2)
```
 

 
 
````{r} 
# - `high_risk_hulls` - made from all high risk sites with `governorate_name` of `Hajjah` or `Marib` except site_id `YE1712_0643` which was clearly outside
# - `gov_hulls` made from all high risk sites falling in the 2 governorates after spatial joining with the admin
# - `district_hull` made from all high risk sies falling the the 3 districts (Abs, Marib, Marib City) based on spatial join with admin
 
 gov_hulls <- c("Hajjah", "Ma'rib") %>%
    map(
      \(gov){
        pts_sf <- high_risk_db_sp %>%
          filter(
            adm1_en == gov,
          ) %>%
          st_as_sf(coords = c("longitude", "latitude"), crs = 4326)
        st_convex_hull(st_union(pts_sf)) %>%
          st_as_sf() %>%
          mutate(governorate_name = gov) %>%
          rename(geometry = x)
      }
    ) %>%
    bind_rows()

 district_hulls <- c("Abs", "Ma'rib") %>%
    map(
      \(district){
        pts_sf <- high_risk_db_sp %>%
          filter(
             str_detect(adm2_en,district) # this will catch Ma'rib and Ma'rib City
            
          ) %>%
          st_as_sf(coords = c("longitude", "latitude"), crs = 4326)
        st_convex_hull(st_union(pts_sf)) %>%
          st_as_sf() %>%
          mutate(dist = district) %>%
          rename(geometry = x)
      }
    ) %>%
    bind_rows()
```



## Current AOI vs AOI defined from 3 Districts
-Let's compare the difference:  <span style="color: orange;">orange - from 3 districts</span>,<span style="color: green;">previous aoi </span>
```{r}
leaflet() %>% 
    addTiles() %>% 
    addPolygons(data=adm_sf$yem_admbnda_adm1_govyem_cso_20191002,fill=NA,fillColor = NA) %>%
    addPolygons(data=adm_sf$yem_admbnda_adm2_govyem_cso_20191002,color="grey",fill=NA,fillColor = NA) %>% 
    addPolygons(data= district_hulls,color="orange", fillColor =NA,fillOpacity = 0) %>%
    addPolygons(data= high_risk_hulls,color="green", fillColor = NA,fillOpacity = 0) 

library(rgee)
ee_Initialize()
chirps_link <- "UCSB-CHG/CHIRPS/DAILY"
chirps <- ee$ImageCollection(chirps_link)
july_rain <- as_tidyee(chirps$filterDate("2021-07-01", "2021-07-22"))

july_sum_ee_img <- chirps$filterDate("2021-07-01", "2021-07-22")$sum()
july_sum<- july_rain %>% 
    # group_by(year, month) %>% 
    summarise(
        stat="sum"
    )

dist_hull_ee <- sf_as_ee(high_risk_district_hulls)
gov_hull_ee <- sf_as_ee(high_risk_hulls)

countDict <- ee$Image$reduceRegions( image= july_sum_ee_img,
                        reducer=ee$Reducer$count(),
                        collection= dist_hull_ee,
                        scale=5500
                        )
countDict <- ee$Image$reduceRegions( image= july_sum_ee_img,
                        reducer=ee$Reducer$count(),
                        collection= gov_hull_ee,
                        scale=5500
                        )
countDict$getInfo()

countDictionary = july_sum_ee_img$reduceRegions(
  reducer=ee$Reducer$count(),
  collection= dist_hull_ee,
  scale= 5500,
  maxPixels= 1e9
)
ee_help(ee$Image$reduceRegions)
countDictionary$getInfo()
pix_count <- ee_extract_tidy(july_sum,
                             y = district_hulls,
                             stat = "count",
                             scale = 5500,via = "getInfo")
    
m1 <- Map$addLayer(july21_mean_rain,visParams = list(min=0,max=10),name = "july21_rain")
library(tidyrgee)
ee_extract_tidy(x = july21_sum_rain)
m1+
    leaflet() %>% 
    addPolygons(data= district_hulls,color="orange", fillColor =NA,fillOpacity = 0) 
```

## Note: Current AOI vs AOI defined from Shapefile Boundaries

- It's worth quickly noting that I originally created the gov level AOI using the `governorate_name` in the master list from CCCM. This may or may not have been a good decision. On the one hand it could be influenced by typos, on the other hand gov names associated with sites by CCCM might be more up to date and/or more operationally relevant than the admin boundaries from shapefile
- Let's compare the difference:  <span style="color: orange;">orange - from shapefile</span>,<span style="color: green;">green - from master list name</span>
-slightly different, but nothing inherently wrong

```{r}
tar_load(high_risk_district_hulls)
leaflet() %>% 
    addTiles() %>% 
    addPolygons(data=adm_sf$yem_admbnda_adm1_govyem_cso_20191002,fill=NA,fillColor = NA) %>%
    addPolygons(data=adm_sf$yem_admbnda_adm2_govyem_cso_20191002,color="grey",fill=NA,fillColor = NA) %>% 
    addPolygons(data= gov_hulls,color="orange", fillColor =NA,fillOpacity = 0) %>%
    addPolygons(data= high_risk_hulls,color="green") 
```

## CHIRPS Comparison


```{r}
tar_load(zonal_stats_high_risk_district_hull)
new_aoi_chirps_rps_full_record <- zonal_stats_high_risk_district_hull$roll3 %>% 
    split(.$governorate_name) %>% 
    imap_dfr(\(df_temp,nm){
        # cat(nm,"/n")
        return_period_level_tibble(df =df_temp,
                                   date = "date",
                                   value = "mean",
                                   rp_year = c(2,3,4,5) ) %>% 
            mutate(governorate_name = nm) 
    }
    )
tar_load(zonal_stats_high_risk_hull)
old_aoi_chirps_rps_full_record <- zonal_stats_high_risk_hull$roll3 %>% 
    split(.$governorate_name) %>% 
    imap_dfr(\(df_temp,nm){
        # cat(nm,"/n")
        return_period_level_tibble(df =df_temp,
                                   date = "date",
                                   value = "mean",
                                   rp_year = c(2,3,4,5) ) %>% 
            mutate(governorate_name = nm) 
    }
    )
```


```{r, results="asis"}

old_aoi_chirps_rps_full_record %>% 
    filter(RP==2) %>% 
    gt() %>%
    fmt_number(columns =everything(),
               decimals = 1) %>% 
    fmt_number(columns =RP,
               decimals = 0) %>% 
    tab_header(title = "2 year RP level CHIRPS (1988-2022) by governorate",
               subtitle ="All high risk sites in Hajjah and Ma'rib governorates" ) %>% 
    opt_align_table_header(align = c("left", "center", "right"))
```

```{r, results="asis"}
new_aoi_chirps_rps_full_record %>% 
    filter(RP==2) %>% 
    gt() %>%
    fmt_number(columns =everything(),
               decimals = 1) %>% 
    fmt_number(columns =RP,
               decimals = 0) %>% 
     tab_header(title = "2 year RP level CHIRPS (1988-2022) by governorate",
               subtitle ="All high risk sites in Abs (Hajjah) and Ma'rib/Ma'rib City (Ma'rib)" ) %>% 
    opt_align_table_header(align = c("left", "center", "right"))
```


```{r}
tar_load(gefs_zonal_districts_rolled)
tar_load(gefs_zonal_rolled)

new_gefs_rps<- gefs_zonal_districts_rolled %>% 
    filter(name =="roll3") %>% 
    filter(!leadtime %in%c(1:2),
           year(date_forecast_predict)!=2023,
           year(date_forecast_predict)!=2020,
           ) %>% 
    mutate(
        gov_lt = paste0(governorate_name,"_",leadtime)
    ) %>% 
    split(.$gov_lt) %>% 
    imap_dfr(\(df_temp,nm){
        # cat(nm,"/n")
        return_period_level_tibble(df =df_temp,
                                   date = "date_forecast_predict",
                                   value = "value",
                                   rp_year = c(2,3,4,5) ) %>% 
            mutate(gov_lt = nm) %>% 
            separate(col = gov_lt,into = c("governorate_name","leadtime"))
    }
    )
old_gefs_rps<- gefs_zonal_rolled %>% 
    filter(name =="roll3") %>% 
    filter(!leadtime %in%c(1:2),
           year(date_forecast_predict)!=2023,
           year(date_forecast_predict)!=2020,
           ) %>% 
    mutate(
        gov_lt = paste0(governorate_name,"_",leadtime)
    ) %>% 
    split(.$gov_lt) %>% 
    imap_dfr(\(df_temp,nm){
        # cat(nm,"/n")
        return_period_level_tibble(df =df_temp,
                                   date = "date_forecast_predict",
                                   value = "value",
                                   rp_year = c(2,3,4,5) ) %>% 
            mutate(gov_lt = nm) %>% 
            separate(col = gov_lt,into = c("governorate_name","leadtime"))
    }
    )
```


```{r}
# get chirps RP for each area
new_chirps_2000_2022_rp <- zonal_stats_high_risk_district_hull$roll3 %>% 
    filter(year(date)%in% c(2000:2022)) %>% 
    filter(year(date)!=2021) %>% 
    split(.$governorate_name) %>% 
    imap_dfr(\(df_temp,nm){
        # cat(nm,"/n")
        return_period_level_tibble(df =df_temp,
                                   date = "date",
                                   value = "mean",
                                   rp_year = c(2,3,4,5) ) %>% 
            mutate(governorate_name = nm) 
    }
    )

old_chirps_2000_2022_rp <- zonal_stats_high_risk_hull$roll3 %>% 
    filter(year(date)%in% c(2000:2022)) %>% 
    filter(year(date)!=2021) %>% 
    split(.$governorate_name) %>% 
    imap_dfr(\(df_temp,nm){
        # cat(nm,"/n")
        return_period_level_tibble(df =df_temp,
                                   date = "date",
                                   value = "mean",
                                   rp_year = c(2,3,4,5) ) %>% 
            mutate(governorate_name = nm) 
    }
    )
```


```{r, results="asis"}
# filter to just 2 year return period
new_chirps_2000_2022_rp %>% 
    filter(RP==2) %>% 
     gt() %>%
    fmt_number(columns =everything(),
               decimals = 1) %>% 
    fmt_number(columns =RP,
               decimals = 0) %>% 
     tab_header(title = "2 year RP level CHIRPS (2000-2022) by governorate",
               subtitle ="All high risk sites in Abs (Hajjah) and Ma'rib/Ma'rib City (Ma'rib)" ) %>% 
    opt_align_table_header(align = c("left", "center", "right")) 
```


```{r,results="asis"}
old_chirps_2000_2022_rp %>% 
    filter(RP==2) %>% 
     gt() %>%
    fmt_number(columns =everything(),
               decimals = 1) %>% 
    fmt_number(columns =RP,
               decimals = 0) %>% 
     tab_header(title = "2 year RP level CHIRPS (2000-2022) by governorate",
               subtitle ="All high risk sites in Hajjah & Ma'rib governorates" ) %>% 
    opt_align_table_header(align = c("left", "center", "right"))
```

```{r, results="asis"}
old_gefs_mean_rp <- old_gefs_rps %>% 
    filter(
        RP==2,
        leadtime %in% c(3:7)
    ) %>% 
    group_by(governorate_name) %>% 
    summarise(
        `previous 2 year RP` = mean(estimate)
    )
new_gefs_mean_rp <- new_gefs_rps %>% 
    filter(
        RP==2,
        leadtime %in% c(3:7)
    ) %>% 
    group_by(governorate_name) %>% 
    summarise(
        `new 2 year RP` = mean(estimate)
    )


left_join(old_gefs_mean_rp,new_gefs_mean_rp) %>% 
    gt() %>%
    fmt_number(columns =everything(),
               decimals = 1) %>% 
     tab_header(title = "2 year RP level CHIRPS-GEFS (2000-2022) by governorate",
               subtitle ="Old method (all gov) vs new method (3 districts)" ) %>% 
    opt_align_table_header(align = c("left", "center", "right"))


```

## Note - Targets Added

analysis above was added as targets so they are easily accessible:

1. a target for convex hull of Abs high risk sites:
    func:   `gen_district_hulls()` 
    target:`high_risk_district_hulls`
2. a target for chirps extraction of high-risk sites. 
    target chirps: `zonal_stats_high_risk_district_hull`
3. a target for chirps-gefs extraction.
    target chirps-gefs : `gefs_zonal_districts_rolled`
