---
title: "Hydrobasins AOI"
output: 
    bookdown::html_document2:
        toc: true
        toc_float: true
        toc_depth: 4
date: "2023-06-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  eval = TRUE,
  results = "asis"
)
```

## Intro

After narrowing down the AOI to the convex hull of just the High Risk Sites in 3 districts we want to explore what it would look like to expand the AOI to:

1. avoid edge effects - sites at the corners/edges of the polyons
2. Incorporate some rainfall in the area surrounding the sitess

**Take away:**

- Basins look promising, but the sizes of **levels 7,8, 9** are not a perfect fit to not overly complicate the project at this phase.
- Therefore a simple buffer seems like a good option



```{r libsDat}
library(rgee)
library(targets)
library(sf)
library(leaflet)
library(tidyverse)
ee_Initialize()
tar_load(adm_sf) # admin units (list of sf features)
tar_load(high_risk_district_hulls)
tar_load(high_risk_site_pts)





aoi_site_points_dist <- high_risk_site_pts %>% 
    st_join(
        adm_sf$yem_admbnda_adm2_govyem_cso_20191002
    ) %>% 
    filter(adm2_en %in% c(
            "Abs",
            "Ma'rib",
            "Ma'rib City"
        )
    )
                        
                    
ee_bbox <- st_bbox(adm_sf$yem_admbnda_adm0_govyem_cso_20191002) %>% 
    st_as_sfc() %>% 
    sf_as_ee()
ee_bbox <- st_bbox(high_risk_district_hulls) %>% 
    st_as_sfc() %>% 
    sf_as_ee()

bas7<- ee$FeatureCollection("WWF/HydroATLAS/v1/Basins/level07")$
    filterBounds(ee_bbox)
bas8<- ee$FeatureCollection("WWF/HydroATLAS/v1/Basins/level08")$
    filterBounds(ee_bbox)
bas9<- ee$FeatureCollection("WWF/HydroATLAS/v1/Basins/level09")$
    filterBounds(ee_bbox)
riv <- ee$FeatureCollection("WWF/HydroSHEDS/v1/FreeFlowingRivers")$
    filterBounds(ee_bbox)
```

## Basin Level 7

Somewhat decent for Hajjah, but a bit hesitant about aggregating the coastal flat and downstream of the sites. Level 7 looks to big for Ma'rib. Also should consider removing that far east point.

```{r}
Map$centerObject(eeObject = ee_bbox,zoom = 8)

m7 <- Map$addLayer(bas7,
                  visParams = list(color="purple"),
                  name = "Basin Level 7")+
    Map$addLayer(riv, visParams = list(color="blue"),name ="Rivers")

m7 %>% 
    addPolygons(data=high_risk_district_hulls ,color = "red",fill = "red") %>% 
    addPolygons(data=adm_sf$yem_admbnda_adm1_govyem_cso_20191002,color = "orange",fillOpacity = 0) %>% 
    addPolygons(data=aoi_buff5k,color = "green",fillOpacity = 0) %>% 
    addPolygons(data=aoi_buff10k,color = "darkgreen",fillOpacity = 0) %>% 
    addMarkers(data= aoi_site_points_dist,label=~site_id)


```

## Basin Level 8

A bit better for Ma'rib, still pretty large.

```{r}
Map$centerObject(eeObject = ee_bbox,zoom = 8)
m8 <- Map$addLayer(bas8,
                  visParams = list(color="purple"),
                  name = "Basin level 8")+
    Map$addLayer(riv, visParams = list(color="blue"),name ="Rivers")

m8 %>% 
    addPolygons(data=high_risk_district_hulls ,color = "red",fill = "red") %>% 
    addPolygons(data=adm_sf$yem_admbnda_adm1_govyem_cso_20191002,color = "orange",fillOpacity = 0) %>% 
    addMarkers(data= aoi_site_points_dist)
```

## Basin Level 9

- **Ma'rib:** Could potentially remove the far east point, and aggregate the 2 basin 9 levels that the rest intersect with.
- **Hajjah:** basins look a bit small and still extend to coast.

```{r}
Map$centerObject(eeObject = ee_bbox,zoom = 8)
m9 <- Map$addLayer(bas9,visParams = list(color="purple"),name = "Basin Level 9")+
    Map$addLayer(riv, visParams = list(color="blue"),name ="Rivers")

m9 %>% 
    addPolygons(data=high_risk_district_hulls ,color = "red",fill = "red") %>% 
    addPolygons(data=adm_sf$yem_admbnda_adm1_govyem_cso_20191002,
                color = "orange",
                fillOpacity = 0) %>% 
    addMarkers(data= aoi_site_points_dist)
```

## Buffer Option

```{r}
tar_load(high_risk_district_hulls_buff)
m7 %>% 
    addPolygons(data=high_risk_district_hulls ,color = "red",fill = "red") %>% 
    addPolygons(data=adm_sf$yem_admbnda_adm1_govyem_cso_20191002,color = "orange",fillOpacity = 0) %>% 
    addPolygons(data=high_risk_district_hulls_buff,color = "green",fillOpacity = 0) 
```
