---
title: "email"
output: blastula::blastula_email
---

<style type = "text/css">

h1, h2, h3 {
  font-family: Arvo, "Helvetica Neue", Helvetica, Arial, sans-serif;
  font-weight: normal
}

p {
  font-family: "Source Sans Pro", "Helvetica Neue", Helvetica, Arial, sans-serif
}

</style>
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fontfamily = "Sans Source Pro",
  fig.cap = ""#,
  # fig.path = file.path("figure", paste0(flag_source, "_", flag_type), "")
)

library(dplyr)
library(ggplot2)
library(glue)
library(gghdx)
library(readr)
library(googledrive)
library(blastula)
library(here)
gghdx()

source(here::here("src/email/email_utils.R"))

thresholds = list(Hajjah = 19.8, Marib=30.6)
```

```{r, class.output = "banner"}

banner_drib <- drive_dribble %>% 
    filter(name == "centre_banner.png")

drive_download(banner_drib, path = banner_path <- tempfile(fileext = ".png"))

add_image(
  file = banner_path,
  alt = "Centre for Humanitarian Data banner",
  align = "left",
  width = "25%"
)

```



```{r}
# shouldn't need to load in gdrive creds when using `envir= parent.frame()`
# zonal_drive <- drive_get(id="1lBZfglOjARZTkVuqyP08O96iW8jimzLu")
# drive_download(zonal_drive, path = f <- tempfile(fileext = ".csv"))
# chirps_gefs_zonal <- read_csv(f)
chirps_gefs_processed <- chirps_gefs_zonal %>% 
    mutate(
        date_predict = date + (leadtime-1)
    ) %>% 
    group_by(governorate_name) %>% 
    arrange(governorate_name,date,leadtime) %>% 
    rename(daily_mm="mean") %>% 
    mutate(
        cumu_3day_mm = zoo::rollsum(x=daily_mm,
                                    k = 3,
                                    fill = NA,
                                    align = "right")
    ) %>% 
    ungroup()

# ck <- chirps_gefs_processed %>% 
#     mutate(
#         cumu_3day_mm =ifelse(date_predict>="2023-05-25",14,cumu_3day_mm)
#     )

trig_status_df <- thresholds %>% 
    imap_dfr(\(num,gov){
        # by gov filter rows > thresh
        df_gte <- chirps_gefs_processed %>% 
            # for txt later
            mutate(
                gov_amt = glue("{governorate_name} ({cumu_3day_mm} mm)")
                ) %>% 
            filter(
                governorate_name==gov &
                    cumu_3day_mm>=num,
            ) 
        if(nrow(df_gte)>0){
            # if any rows  make sure we get min leadtime
          ret <- df_gte %>% 
            filter(
                leadtime ==min(leadtime)
            )
        } else{
            ret <- df_gte
        }
    }
    ) #%>% 
    # rm record w/ out any vals gte thresh
    # discard(.p = \(df_temp) nrow(df_temp)==0)


both_txt <- "based on the 7-day forecast from CHIRPS-GEFS."

if(nrow(trig_status_df)==0){
    pre_txt <- "**has not been activated**"
    where_txt <- ""
    spec_txt <- ""
}
if(nrow(trig_status_df)>0){
    pre_txt <- "**has been activated in**"
    where_txt <-  glue_collapse(x = trig_status_df$governorate_name,sep = " & ")
    when_txt <-  glue_collapse(x = trig_status_df$date_predict,sep = " & ") 
    amt_txt <-  glue_collapse(x = trig_status_df$cumu_3day_mm,sep = " & ") 
    gov_amt_txt <-  glue_collapse(x = trig_status_df$gov_amt,sep = " & ") 
    spec_txt <- glue("The forecast predicts trigger thresholds to be crossed on {when_txt} for {gov_amt_txt}{ifelse(nrow(trig_status_df)>1,', respectively.', '.')}")
    
}

desc_txt <- glue_collapse(c(pre_txt,where_txt,both_txt,spec_txt),sep=" ")

```

# Yemen Rainfall Forecast Monitoring

## Anticipatory Action Flooding Trigger Status - `r format(Sys.Date(),"%d %B %Y")`

The AA Flooding Trigger for Yemen `r desc_txt`

The AA Flooding Trigger is based on 3-day cumulative rainfall thresholds for our Areas of Interest (AOI) withing Marib and Hajjah governorates in Yemen. The selected thresholds are based on a 2-year return period event and differ for each of the two locations. In Hajjah and Marib the current thresholds are set to **19.8 mm** and **30.6 mm**, respectively.



```{r makeplot}
forecast_date<- min(chirps_gefs_processed$date_predict)
date_7day <-  chirps_gefs_processed %>% 
    filter(leadtime ==7) %>% 
    pull(date_predict) %>% 
    unique()

trigger_lines <- data.frame(
    governorate_name=c("Hajjah","Marib"),
    threshold = c(19.8,30.6),
    label_place_x = forecast_date+1
)

subtitle <- glue("Forecast date: {forecast_date}\n")
                 
                 
p <- chirps_gefs_processed %>% 
    mutate(
        status= ifelse(leadtime %in% c(1:7),"Trigger Window (7 days)","Pre-Alert"),
        status=factor(status, levels=c("Trigger Window (7 days)","Pre-Alert"))
    ) %>% 
    ggplot(aes(x= date_predict,y = cumu_3day_mm))+
    geom_line(aes(color=status))+
    geom_point(aes(color=status))+
    scale_color_manual(values = c("#347ad6ff","orange"))+
    scale_x_date(date_breaks = "day",date_labels = "%y-%m-%d")+
    geom_hline(data = trigger_lines,
               aes(yintercept=threshold),color="red",
               linetype="dashed"
               )+
    geom_text(data= trigger_lines,
              aes(x=label_place_x, y=threshold+1.5,
                  label=glue("{governorate_name} threshold: {threshold}")
                  ),color="red",size=6)+
    facet_wrap(~governorate_name,ncol=1,scales="free_y")+
    labs(title = "Yemen 10 Day Forecast (CHIRPS-GEFS)",
         subtitle = subtitle,
         y= "3 Day Rainfall (mm)"
         )+
    theme_hdx()+
    theme(
        
        title = element_text(size= 24),
        axis.title = element_text(size=18),
        axis.text.x = element_text(angle=90),
        axis.title.x =element_blank(),
        # panel.spacing = unit(5, "lines"),
        panel.border = element_rect(color = "grey",fill = NA),
        legend.title = element_blank()
    )
```

```{r, results ="asis"}
add_ggplot_custom(
  plot_object = p,
  alt = "Alert plot",
  height = 6.5,
  width = 6,
  html_width = 1000
)

```



----

```{r, eval=F,results = "asis"}

pwalk(
  .l = flags_specific %>%
    select(
      iso3,
      country,
      start_date,
      end_date,
      message,
      url,
      summary_experimental
    ),
  .f = \(
    iso3,
    country,
    start_date,
    end_date,
    message,
    url,
    summary_experimental
  ) {
    knitr::knit_child(
      "alert_section.Rmd",
      envir = environment(),
      quiet = TRUE
    ) |>
      cat()
  }
)
```

## Contact

Contact the OCHA Centre for Humanitarian Data via Leonardo Milano, Team Lead
for Data Science at leonardo.milano@un.org with any questions or feedback.

----

```{r}
logo_drib <- drive_dribble %>% 
    filter(name == "ocha_logo_wide.png")
drive_download(logo_drib, path = logo_path <- tempfile(fileext = ".png"))

add_image(
  file = logo_path,
  align = "center",
  width = "25%"
)
```

<p>
<center>

<b> OCHA Centre For Humanitarian Data </b>

Fluwelen Burgwal 58 | 2511 CJ The Hague | The Netherlands

</center>
</p>
