---
title: "email"
output: blastula::blastula_email
---

<style type = "text/css">

h1, h2, h3 {
  font-family: Arvo, "Helvetica Neue", Helvetica, Arial, sans-serif;
  font-weight: normal
}

p {
  font-family: "Source Sans Pro", "Helvetica Neue", Helvetica, Arial, sans-serif
}

</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fontfamily = "Sans Source Pro",
  fig.cap = "" # ,
  # fig.path = file.path("figure", paste0(flag_source, "_", flag_type), "")
)

library(dplyr)
library(ggplot2)
library(glue)
library(gghdx)
library(readr)
library(googledrive)
library(blastula)
library(here)
gghdx()

source(here::here("src/email/email_utils.R"))

# adjusted thresholds based on chirp-gefs RP 2 mean(lt3:7) `10_CHIRPS_GEFS.Rmd`
thresholds <- list(Hajjah = 23.8, Marib = 21.8)
```

```{r, class.output = "banner"}
banner_drib <- drive_dribble %>%
  filter(name == "centre_banner.png")

drive_download(banner_drib, path = banner_path <- tempfile(fileext = ".png"))

add_image(
  file = banner_path,
  alt = "Centre for Humanitarian Data banner",
  align = "left",
  width = "25%"
)
```



```{r}
chirps_gefs_processed <- chirps_gefs_zonal %>%
  mutate(
    date_predict = date + (leadtime - 1)
  ) %>%
  group_by(governorate_name) %>%
  arrange(governorate_name, date, leadtime) %>%
  rename(daily_mm = "mean") %>%
  mutate(
    cumu_3day_mm = zoo::rollsum(
      x = daily_mm,
      k = 3,
      fill = NA,
      align = "right"
    )
  ) %>%
  ungroup()

trig_status_df <- thresholds %>%
  imap_dfr(\(num, gov){
    # by gov filter rows > thresh
    df_gte <- chirps_gefs_processed %>%
      # for txt later
      mutate(
        gov_amt = glue("{governorate_name} ({cumu_3day_mm} mm)")
      ) %>%
      filter(
        governorate_name == gov &
          cumu_3day_mm >= num,
      )
    if (nrow(df_gte) > 0) {
      # if any rows  make sure we get min leadtime
      ret <- df_gte %>%
        filter(
          leadtime == min(leadtime)
        )
    } else {
      ret <- df_gte
    }
  }) # %>%
# rm record w/ out any vals gte thresh
# discard(.p = \(df_temp) nrow(df_temp)==0)


# both_txt <- "based on the 7-day forecast from CHIRPS-GEFS."

if (nrow(trig_status_df) == 0) {
  pre_txt <- "**has not been activated.**  The forecast does not predict trigger thresholds to be crossed in the next 7 days"
  where_txt <- ""
  spec_txt <- ""
}
if (nrow(trig_status_df) > 0) {
  pre_txt <- "**has been activated in**"
  where_txt <- glue(glue_collapse(x = trig_status_df$governorate_name, sep = " & "),".")
  when_txt <- glue_collapse(x = trig_status_df$date_predict, sep = " & ")

  amt_txt <- glue_collapse(x = trig_status_df$cumu_3day_mm, sep = " & ")
  gov_amt_txt <- glue_collapse(x = trig_status_df$gov_amt, sep = " & ")
  
  same_date_cond <- length(unique(trig_status_df$date_predict))==1
  both_gov_cond <- length(unique(trig_status_df$governorate_name))==2
  
  spec_txt <- glue("The forecast predicts trigger thresholds to be crossed on {ifelse(same_date_cond,unique(as.character(trig_status_df$date_predict)),when_txt)} in {gov_amt_txt} {ifelse(!same_date_cond,', respectively', '')}")
}

desc_txt <- str_trim(glue_collapse(c(pre_txt, where_txt, spec_txt), sep = " "))
```

# Yemen Rainfall Forecast Monitoring

## Anticipatory Action Flooding Trigger Status - `r format(Sys.Date(),"%d %B %Y")`

The AA Flooding Trigger for Yemen `r desc_txt`.

The AA Flooding Trigger is based on 3-day cumulative rainfall thresholds for our Areas of Interest (AOI) within Marib and Hajjah governorates in Yemen. The selected thresholds are based on a 2-year return period event and differ for each of the two locations. In Hajjah and Marib the current thresholds are set to **19.8 mm** and **30.6 mm**, respectively.

```{r makeplot}
forecast_date <- min(chirps_gefs_processed$date_predict)

trigger_lines <- data.frame(
  governorate_name = c("Hajjah", "Marib"),
  threshold = c(19.8, 30.6),
  label_place_x = forecast_date + 1
)

subtitle <- glue("Forecast date: {forecast_date}\n")

p <- chirps_gefs_processed %>%
  mutate(
    status = ifelse(leadtime %in% c(1:7), "Trigger Window (7 days)", "Pre-Alert"),
    status = factor(status, levels = c("Trigger Window (7 days)", "Pre-Alert"))
  ) %>%
  ggplot(aes(x = date_predict, y = cumu_3day_mm)) +
  geom_line(aes(color = status)) +
  geom_point(aes(color = status)) +
  scale_color_manual(values = c("#347ad6ff", "orange")) +
  scale_x_date(date_breaks = "day", date_labels = "%y-%m-%d") +
  geom_hline(
    data = trigger_lines,
    aes(yintercept = threshold), color = "red",
    linetype = "dashed"
  ) +
  geom_text(
    data = trigger_lines,
    aes(
      x = label_place_x, y = threshold + 1.5,
      label = glue("{governorate_name} threshold: {threshold}")
    ), color = "red", size = 4
  ) +
  facet_wrap(~governorate_name, ncol = 1, scales = "free_y") +
  labs(
    title = "Yemen 10 Day Forecast (CHIRPS-GEFS)",
    subtitle = subtitle,
    y = "3 Day Rainfall (mm)"
  ) +
  theme_hdx() +
  theme(
    title = element_text(size = 14),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(angle = 90),
    axis.title.x = element_blank(),
    legend.text = element_text(size=12),
    # panel.spacing = unit(5, "lines"),
    panel.border = element_rect(color = "grey", fill = NA),
    legend.title = element_blank()
  )
```

```{r, results ="asis"}
add_ggplot_custom(
  plot_object = p,
  alt = "Alert plot",
  height = 5,
  width = 4.5,
  html_width = 750
)
```

**Forecast data source:** [CHIRPS-GEFS](https://chc.ucsb.edu/data/chirps-gefs) Precipitation Forecasts https://doi.org/10.15780/G2PH2M (2021). Data was accessed on `r Sys.Date()`. 

----

## Contact

Contact the OCHA Centre for Humanitarian Data via Leonardo Milano, Team Lead
for Data Science at leonardo.milano@un.org with any questions or feedback.

----

```{r}
logo_drib <- drive_dribble %>%
  filter(name == "ocha_logo_wide.png")
drive_download(logo_drib, path = logo_path <- tempfile(fileext = ".png"))

add_image(
  file = logo_path,
  align = "center",
  width = "25%"
)
```

<p>
<center>

<b> OCHA Centre For Humanitarian Data </b>

Fluwelen Burgwal 58 | 2511 CJ The Hague | The Netherlands

</center>
</p>
