---
title: "Yemen Impact Flooding Data"
output: html_document
date: "2023-02-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Yemen Impact Flooding Data

This doc is to mainly understand the various impact data source that could potentially be used in our analysis:

So far we have identified the following:

- EMDAT
- Desinventar
- REACH/CCCM
- CCCM
    + CCCM flood incident reporting (2021 & 2022)
    + CCCM Site master list

This doc is purely exploratory. In it's current state it is meant to be run line by line, but can be refactored for knitting if necessary

**TODO** add YEM CODs to AA_DATA_DIR? **NOTE:** won't be fully runnable by all team members until this is done


```{r yem_impact_libs}
library(tidyverse)
library(here)
library(readxl)
library(janitor)
library(sf)
library(lubridate)
# library(tidytext)
library(snakecase)
library(gghdx)
library(glue)
library(sysfonts)

font_add_google("Source Sans Pro")
showtext::showtext_auto()

```


## Loading data
```{r}

emdat <- read_excel( file.path(Sys.getenv("AA_DATA_DIR"),"public","exploration","yem",
                               "yem_emdat_flood_natural_events.xlsx" ),skip = 6) %>% 
    clean_names()

desinv <- read_excel(file.path(Sys.getenv("AA_DATA_DIR"),"public","exploration","yem","yem_desinventar_flood_events.xlsx") ) %>% 
    clean_names()
desinv %>% 
    glimpse()
```


Seems neither has any gps locations - emailed CCCCM colleagues Feb 15 to ask about this
```{r}
cccm_master_fp <- file.path(Sys.getenv("AA_DATA_DIR"),"public","exploration","yem","Yemen -  CCCM Cluster -December IDP Sites Master List - 20221231.xlsx")
cccm_fp <- file.path(Sys.getenv("AA_DATA_DIR"),"public","exploration","yem","Yemen -  CCCM Cluster -December ML - Flooding available data(18).xlsx")

# grab sheet names
cccm_sheets <- cccm_fp %>% 
    excel_sheets()

cccm_master_sheets <-  cccm_master_fp %>% 
    excel_sheets()

# read in alll tabs of master list
cccm_master_skips <- c(1,1,0,3,1,1)
cccm_master_dfs <- cccm_master_sheets %>% 
    map2(
        .y=cccm_master_skips,
        ~read_excel(cccm_master_fp,sheet=.x,skip = .y)
    ) %>% 
    set_names(cccm_master_sheets) 

# query for gps/coordinate info
data.frame(
    cccm_colnames=cccm_master_dfs %>% 
    map(~colnames(.x)) %>% 
        unlist() %>% 
        to_snake_case()
) %>% 
    filter(str_detect(cccm_colnames,"loc|gps|coord|geo"))


skip_vals <- c(1,rep(0,3))

cccm_dfs <- cccm_sheets %>% 
    map2(.y = skip_vals,
        ~read_excel(cccm_fp,sheet=.x,skip = .y) %>% 
            clean_names()
    ) %>% 
    set_names(cccm_sheets) 

data.frame(
    cccm_colnames=cccm_dfs %>% 
    map(~colnames(.x)) %>% unlist() %>% 
        to_snake_case()
) %>% 
    filter(str_detect(cccm_colnames,"location|gps|coordinate|latit|long"))
```


Should load in a YEM admin COD, but not yet in AA_DATA_DIR.. for now I will just use my db, but need to replace this
```{r}

####################################################
### will replace this seciton with AA_DATA_DIR #####
#####################################################
con <- DBI::dbConnect(RPostgres::Postgres(),
                      host= "localhost",
                      dbname = "global_gdb",
                      user      = "postgres",
                      # password      = db_info$drc_chol_db_pw,
                      port     = 5432)

adm3 <- st_read(con, "yem_adm3")

```


## Start Exploring
 
### CCCCM 

```{r}
cccm_dfs$`ML DEC2022 - Flooding frequency` %>% 
    skimr::skim()

cccm_dfs$`FLOOD REPORT IN SITES - Dec2022` %>% 
    skimr::skim()

cccm_dfs$Sheet1

cccm_dfs$Sheet5

cccm_master_dfs$`CCCM IDP SITES`$`Location/Village Name` %>% tabyl()
cccm_master_dfs$`CCCM IDP SITES` %>% 
    skimr::skim()
cccm_master_dfs$GRAPH

cm_flood <- cccm_dfs$`FLOOD REPORT IN SITES - Dec2022` 

cm_flood <- cm_flood %>% select(date_of_episode,
                    num_init_hhs =initial_figures_h_hs,
                    num_verified_hhs = verified_figures_h_hs,
                    num_shelters_affefcted= how_many_shelters_were_affected,
                    num_hhs_lost_documentation=if_yes_how_many_h_hs_did_they_lose_their_personal_legal_documentations_due_to_the_flooding,
                    everything())
cm_flood %>% glimpse()
```



Make CCCM plots

checkout flood reports on a timeline

```{r}
p_cccm_total_hhs_v <- cm_flood %>% 
    group_by(date_of_episode) %>% 
    summarise(
        across(starts_with("num_"),~sum(.x,na.rm=T))
        ) %>% 
    ggplot(aes(x=ymd(date_of_episode), y= num_verified_hhs)) +
    geom_point()+
    scale_x_date(date_breaks = "month",
                 date_labels = "%b-%y"
                 )+
    labs(title= "CCCM Flood Report Data: Number Verified HHs Affected",subtitle = "Yemen (2021-2022)")+
    theme_hdx()+
    theme(
        axis.text.x = element_text(angle=90),
        axis.title = element_blank()
          )
p_cccm_total_hhs_v
```


by governorate - this plot seem potentially useful

```{r}
p_cccm_total_hhs_v_gov <- cm_flood %>% 
    group_by(date_of_episode,governorate_name) %>% 
    summarise(
        across(starts_with("num_"),~sum(.x,na.rm=T)),.groups = "drop"
        ) %>% 
    ggplot(aes(x=ymd(date_of_episode), y= num_verified_hhs, group=governorate_name,
               color=governorate_name)) +
    geom_point()+
    geom_line()+
    scale_x_date(date_breaks = "month",
                 date_labels = "%b-%y"
                 )+
    scale_y_log10_hdx()+
    # scale()+
    labs(title= "CCCM Flood Report Data: Number Verified HHs Affected by Governorate",subtitle = "Yemen (2021-2022)")+
    facet_wrap(~governorate_name)+
    theme_hdx()+
    theme(
        legend.position = "none",
        axis.text.x = element_text(angle=90),
        axis.title = element_blank()
          )

p_cccm_total_hhs_v_gov

```

Also by governorate

```{r}
cm_flood %>% 
    group_by(governorate_name) %>%
    summarise(
        num_records=n()
        ) %>% 
    ggplot(aes(x= reorder(governorate_name,num_records), y= num_records))+
    geom_col()+
    geom_text(aes(label=num_records),hjust=-1.2)+
    scale_y_continuous(breaks = seq(0,175, by=25), limits = c(0,150))+
    coord_flip()+
    labs(title= "CCCM Flood Report Data: Number Reports by Governorate",
         subtitle = "Yemen (2021-2022)")+
    
    
    theme_hdx()+
    theme(
        axis.text.x = element_text(angle=90),
        axis.title = element_blank()
          )
```



## EM-DAT

Let's look at some summary stats and completeness of data to what might be useable:

 - year - great (100% completeness)
 - month decent (95%)
 - deaths (80%)
 - total affected (76 %)
 
 - coordinates unusable (0%)
 - river basin has (0%)
 
```{r}
skimr::skim(emdat)
```

let's breakd down the types and subtypes to make sure we are not missing anything... from this table it looks like filtering to disaster
```{r}
emdat %>% count(event_name)
```

break down types further and decide to filter based on `disaster_type` for now
```{r}
emdat %>% count(disaster_subgroup,disaster_type,disaster_subtype,disaster_subsubtype)

emf <- emdat %>% 
    filter(disaster_type=="Flood")
```


after filter let's skim again
```{r}
skimr::skim(emf)
```

unfortunately the admin1 codes do not bear any semblance of admin1 codes in the COD which are 2 digit numbers that follow a "Y"
```{r}
emf %>% 
    count(country,iso,region,location)

emf %>% 
    count(country,iso,region,admin1_code) %>% 
    print(n=nrow(.))

emf %>% 
    tidyr::separate_rows(admin1_code,sep=";") %>% 
    select(admin1_code,everything()) %>% 
    count(admin1_code) %>% print(n=nrow(.))

```


Let's visualized

- number of incidents on a timeline
```{r}

p_number_incidents <- emf %>% 
    count(year) %>%
    mutate(year = as.numeric(year)) %>% 
    complete(year = full_seq(year, period = 1), fill = list(n = NA)) %>% 
    ggplot(aes(x= year,y=n))+
    geom_point()+
    labs(y= "# incidents",title = "EM-DAT: Number of Flood Related Records Per Year",subtitle = "Yemen")+
    theme_hdx()

p_number_incidents
```

same thing, but by type

```{r}


emf %>% 
    count(year,disaster_subtype) %>%
    mutate(year = as.numeric(year),
           # type=fct_explicit_na(disaster_subtype)
           `Flood Type`= ifelse(is.na(disaster_subtype),"Unreported Type",disaster_subtype)
           ) %>% 
    
    complete(year = full_seq(year, period = 1), fill = list(n = NA)) %>% 
    ggplot(aes(x= year,y=n, color=`Flood Type`))+
    geom_point()+
    labs(y= "# incidents",title = "EM-DAT: Number of Flood Related Records Per Year",subtitle = "Yemen")+
    scale_color_discrete(na.translate=FALSE) +
    theme_hdx()
```


summarise affected/deaths by year and year/month to get some sumnary stats to plot

```{r}


emf_impact_yearly <- emf %>% 
   group_by(year) %>% 
    summarise(
        across(c("total_affected","total_deaths"),~sum(.x,na.rm=T))
    )
emf_impact_yrmo <- emf %>% 
   group_by(year,start_month) %>% 
    summarise(
        across(c("total_affected","total_deaths"),~sum(.x,na.rm=T)),.groups = "drop"
    )



emf_impact_type_yearly <- emf %>% 
    mutate(
        `Flood Type`= ifelse(is.na(disaster_subtype),"Unreported Type",disaster_subtype)
    ) %>% 
   group_by(year,`Flood Type`) %>% 
    summarise(
        across(c("total_affected","total_deaths"),~sum(.x,na.rm=T)),.groups = "drop"
    )

p_total_affected <- emf_impact_yearly %>% 
    mutate(year = as.numeric(year)) %>% 
    complete(year = full_seq(year, period = 1), fill = list(n = NA)) %>% 
    ggplot(aes(x= year,y=total_affected))+
    geom_point()+
    scale_y_continuous(labels = scales::comma)+
    labs(y= "# Affected",title = "EM-DAT Flood Data: People Affected",subtitle = "Yemen")+
    theme_hdx()+
    theme(
        axis.title.y = element_blank()
    )

p_total_deaths <- emf_impact_yearly %>% 
    mutate(year = as.numeric(year)) %>% 
    complete(year = full_seq(year, period = 1), fill = list(n = NA)) %>% 
    ggplot(aes(x= year,y=total_deaths))+
    geom_point()+
    scale_y_continuous(labels = scales::comma)+
    labs(y= "# Deaths",title = "EM-DAT Flood Data: Reported Deaths",subtitle = "Yemen")+
    theme_hdx()+
    theme(
        axis.title.y = element_blank()
    )

p_type_affected <- emf_impact_type_yearly %>% 
    mutate(year = as.numeric(year)) %>% 
    complete(year = full_seq(year, period = 1), fill = list(n = NA)) %>% 
    ggplot(aes(x= year,y=total_affected,color=`Flood Type`))+
    geom_point()+
    scale_y_continuous(labels = scales::comma)+
    labs(y= "# Affected",title = "EM-DAT Flood Data: People Affected",subtitle = "Yemen")+
    theme_hdx()+
    theme(
        axis.title.y = element_blank()
    )+
    scale_color_discrete(na.translate=FALSE) 

p_type_death <- emf_impact_type_yearly %>% 
    mutate(year = as.numeric(year)) %>% 
    complete(year = full_seq(year, period = 1), fill = list(n = NA)) %>% 
    ggplot(aes(x= year,y=total_deaths,color=`Flood Type`))+
    geom_point()+
    scale_y_continuous(labels = scales::comma)+
    labs(y= "# Deaths",title = "EM-DAT Flood Data: Reported Deaths",subtitle = "Yemen")+
    theme_hdx()+
    theme(
        axis.title.y = element_blank()
    )+
    scale_color_discrete(na.translate=FALSE) 

```

plot reported deaths/affetcted over time and by type

```{r}

ggpubr::ggarrange(p_total_affected,
                  p_type_affected,
                  p_total_deaths,
                  p_type_death,ncol = 2,nrow=2
                  )

```



Let's look at affected by month to see if there is a pattern. Also let's put all events on a complete timeline and label events with especially high numbers of people affected

```{r}
emf_impact_yrmo <- emf_impact_yrmo %>% 
    mutate(year = as.numeric(year)) %>% 
    complete(
        year = full_seq(year, period = 1), fill = list(n = NA),
             ) %>% 
    complete(
            year, start_month=full_seq(1:12, period =1),fill = list(n = NA)
    ) %>% 
    arrange(year,start_month) %>% 
    mutate(
        month_chr= month(start_month,abbr = T,label=T),
        date = make_date(year= year, month= start_month,day=1)
    )




p_total_affected_by_mo <- emf_impact_yrmo %>% 
    group_by(month_chr) %>% 
    summarise(
        total_affected= sum(total_affected,na.rm=T),
        total_deaths= sum(total_deaths,na.rm=T)
    ) %>% 
    filter(!is.na(month_chr)) %>% 
    # pivot_longer(total_affected:total_deaths,names_to = "death_affected") %>% 
    ggplot(aes(x=month_chr, y =total_affected))+
    scale_y_continuous_hdx(labels=scales::comma)+
    geom_bar(stat="identity")+
    theme_hdx()+
    labs(title = "EM-DAT Flood Data: People Affected By Month",
         subtitle = "Yemen (1972-2022)")+
    theme(
        legend.position = "none",
        axis.title = element_blank()
    )




p_affected_by_date <- emf_impact_yrmo %>%
    mutate(
        aff_lab = if_else(total_affected>5e4,glue("{month_chr} ({str_sub(year(date),3,4)})"),NA_character_)
    ) %>% 
    filter(year(date)<=2023) %>% 
    ggplot(aes(x=date, y=total_affected))+
    geom_point()+
    geom_text(aes(label=aff_lab),vjust=2)+
    scale_x_date(date_breaks = "year",date_labels = "%Y")+
    scale_y_continuous_hdx(labels=scales::comma,limits=(c(0,375e3)))+
    labs(
        title = "EM-DAT Flood Data: People Affected",
         subtitle = "Yemen (1972-2022)")+
    theme_hdx()+
    theme(
        axis.text.x = element_text(angle=90),
        axis.title = element_blank()
    )

ggpubr::ggarrange(p_total_affected_by_mo,
                  p_affected_by_date,
                  ncol = 1,nrow=2
                  )

```

- hmm so I wonder if we can pull precip/spi in and look at those dates...but what about the geographic component
- here is initially data cleaning work to disaggregate the data by admin 1 - this might be useful, but will require some assumptions.

***note:**  if NA shows in admin1 column that means I have not found a suitable admin1 from the master list


```{r}

adm3$admin1Name_en %>% unique() %>% sort()
to_snake_case(adm3$admin1Name_en) %>% unique() %>% sort()


emf %>% 
    tidyr::separate_rows(location,sep=",") %>% 
    mutate(
        loc_snake = to_snake_case(location),
        loc_snake=stringi::stri_trans_general(str = loc_snake, id = "Latin-ASCII"),
        location= trimws(location),
        admin1= case_when(
            str_detect(loc_snake,"san")~ "sana_a",
            str_detect(loc_snake,"sad|saad|sa_ad")~ "sa_dah",
            str_detect(loc_snake,"aby|abin")~ "abyan",
            str_detect(loc_snake,"ade")~ "aden",
            str_detect(loc_snake,"^ib")~ "ibb",
            str_detect(loc_snake,"^ad_dal|^al_dhal|^al_dal|dhale")~ "ad_dali",
            str_detect(loc_snake,"^al_bay")~ "al_bayda",
            str_detect(loc_snake,"^al_hode|^al_hudayd|hodeida|hudayda")~ "al_hodeidah",
            str_detect(loc_snake,"^al_mahr|^al_mahar")~ "al_maharah",
            str_detect(loc_snake,"^al_mahw")~ "al_mahwit",
            str_detect(loc_snake,"^el_jouf|alljawf|al_jawf|aljawf")~ "al_jawf",
            str_detect(loc_snake,"marib|ma_rib|mareb")~ "ma_rib",
            str_detect(loc_snake,"hadram|hadhram")~ "hadramawt",
            str_detect(loc_snake,"dhama")~ "dhamar",
            str_detect(loc_snake,"hajja")~ "hajjah",
            str_detect(loc_snake,"rayma")~ "raymah",
            str_detect(loc_snake,"^lahj|^lah[ei]j")~ "lahj",
            str_detect(loc_snake,"ta[ie]z")~ "ta_iz",
            str_detect(loc_snake,"shabwa")~ "shabwah",
            str_detect(loc_snake,"socotra")~ "socotra",
            str_detect(loc_snake,"amran")~ "amran",
            TRUE~ NA_character_
        )
        ) %>% 
    count(location,admin1) %>% 
    arrange(location) %>% 
    print(n=nrow(.))
```


## Desinventar

let's skim desvintar dta

all data is flood

```{r}
desinv %>% 
    skimr::skim()

desinv %>% 
    count(event)

desinv <- desinv %>% 
    mutate(
        date_fix=janitor::excel_numeric_to_date(as.numeric(date))
    ) 
```



what if we combine the desinventar flood dates with the EM-DAT. Below I add vertical lines for Desinventar flood dates:
```{r}

p_affected_by_date <- emf_impact_yrmo %>%
    mutate(
        aff_lab = if_else(total_affected>5e4,glue("{month_chr} ({str_sub(year(date),3,4)})"),NA_character_)
    ) %>% 
    filter(year(date)<=2023) %>% 
    ggplot(aes(x=date, y=total_affected))+
    geom_point()+
    geom_text(aes(label=aff_lab),vjust=2)+
    scale_x_date(date_breaks = "year",date_labels = "%Y")+
    scale_y_continuous_hdx(labels=scales::comma,limits=(c(0,375e3)))+
    labs(
        title = "EM-DAT Flood Data: People Affected",
         subtitle = "Yemen (1972-2022)")+
    geom_vline(xintercept = desinv %>% filter(!is.na(date_fix)) %>% pull(date_fix))+
    theme_hdx()+
    theme(
        axis.text.x = element_text(angle=90),
        axis.title = element_blank()
    )

```

```{r}
p_affected_by_date
```